---
title: "Visualization of figures"
author: "Sophie"
date: "2026-04-16"
output: html_document
---
Load necessary libraries
```{r}
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(sf)
library(INLA)
library(viridis)
library(scales)
library(lubridate)
library(GHRmodel)
library(lubridate)
library(RColorBrewer)
```
# Scripts for visualizing Main Figure 1
## Figure 1A: visualize hospital locations
```{r}
geocoded_hospitals_dists<-readRDS(file="/home/sbelman/Documents/env_sa_manuscript/input_datasets/germs/geocoded_hospitals_districts.rds" )

# TURN TO AN SF
points_sf <- st_as_sf(geocoded_hospitals_dists, coords = c("long", "lat"), crs = 4326)
# Check the CRS of the shapefile and points
correct<-st_crs(shp)
st_crs(points_sf)
# If needed, transform the CRS of points to match the shapefile
points_sf <- st_transform(points_sf, correct)
color_labels = c("Eastern Cape" = "#008080", "Free State" = "#D75F00", "Gauteng" = "#967bb6" , "KwaZulu-Natal"= "#FFCCCB",
                     "Limpopo" = "#228B22", "Mpumalanga" = "#FFC000", "North West" = "#8B5A2B", "Northern Cape" = "#606060", "Western Cape" = "#1A2D6D" )
# st_crs(points_sf)
pprov<-ggplot(shp)+
  geom_sf()+
  # geom_sf(data = points_sf,  size = 3) +
  geom_sf(data = points_sf, aes(color = factor(province)), size = 3) +
  scale_color_manual(values=color_labels)+
  theme_classic()+
  theme(axis.text = element_text(size=13), legend.text = element_text(size=13), legend.title = element_text(size=13))+
  labs(color="Province")
# pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure1/hosp_pprov.pdf", width = 7,height=5)
# print(pprov)
# dev.off()
```

#### Figure 1B: PLOT DISEASE INCIDENCE
```{r}

shp<-st_read("/home/sbelman/Documents/env_sa_manuscript/input_datasets/shps/gadm41_namematch_ZAF_2.shp")
sa_adm2_weekly<-fread(file="/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm2_weekly.csv")

dis_data <- sa_adm2_weekly %>%
  group_by(district)%>%
  summarise(
    disease = sum(disease, na.rm = TRUE),
    population_size = first(population_size),
    .groups = "drop"
  )

shp_dis <- shp %>%
  left_join(dis_data, by = "district")
# Define nice breaks for the legend
legend_breaks <- c(1, 10, 100, 1000, 10000, 50000, 100000)
# legend_breaks_inc <- c(1, 5, 10, 50, 100, 500, 1000)

# Plot with log-transformed color scale but real value labels
## incidence
displotinc <- ggplot(shp_dis) +
  geom_sf(aes(fill = (disease/population_size)*100000), color = "white", size = 0.1) +
  scale_fill_viridis(trans = "log10",
                     # breaks = legend_breaks_inc, labels = scales::comma(legend_breaks_inc),
                     na.value = "gray90",name = "Disease Incidence\nper 100,000") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold", size = 16),plot.subtitle = element_text(size = 12),legend.position = "right",
    panel.grid = element_blank(),axis.text = element_blank(),axis.title = element_blank(),axis.ticks = element_blank())
## count
displot_count <- ggplot(shp_dis) +
  geom_sf(aes(fill = disease), color = "white", size = 0.1) +
  scale_fill_viridis(trans = "log10",breaks = legend_breaks,labels = scales::comma(legend_breaks),
                     na.value = "gray90",name = "Disease counts") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold", size = 16),plot.subtitle = element_text(size = 12),legend.position = "right",
    panel.grid = element_blank(),axis.text = element_blank(),axis.title = element_blank(),axis.ticks = element_blank())
displot_count+displotinc
    
pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure1/dis_inc.pdf", width = 7,height=5)
print(displotinc)
dev.off()

pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure1/dis_counts.pdf", width = 7,height=5)
print(displot_count)
dev.off()
```
## Figure 1C: Summary of pm2.5 max temp, relative and absolute humidity over time in facet
```{r}
sa_adm2_weekly<-fread(file="/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm2_monthly.csv")
## colored by province
env_data <- sa_adm2_weekly %>%
  group_by(date,NAME_1)%>%
  summarise(
    disease = sum(disease, na.rm=T),
    age_gt65 = sum(age_gt65, na.rm = T),
    pm2p5 = mean(pm2p5, na.rm = TRUE),
    tasmax = mean(tasmax,na.rm = TRUE),
    tas = mean(tas,na.rm = TRUE),
    hurs = mean(hurs,na.rm = TRUE),
    absh = mean(absh,na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(disease, age_gt65, pm2p5, tasmax,"tas", hurs, absh),
    names_to = "variable",
    values_to = "value"
  )
color_labels = c("Eastern_Cape" = "#008080", "Free_State" = "#D75F00", "Gauteng" = "#967bb6" , "KwaZulu-Natal"= "#FFCCCB",
                     "Limpopo" = "#228B22", "Mpumalanga" = "#FFC000", "North_West" = "#8B5A2B", "Northern_Cape" = "#606060", "Western_Cape" = "#1A2D6D" )
env_data$variable <- factor(env_data$variable, levels = c("disease","age_gt65","pm2p5","tasmax","tas","hurs","absh"))

envs <- ggplot(env_data)+
  geom_line(aes(x=date,y=value, group=interaction(variable,NAME_1), color=NAME_1),alpha=0.8)+
  theme_bw()+
  facet_grid(variable~.,scales="free")+
  scale_color_manual(values=color_labels)+
  theme(legend.position = "right",legend.title = element_blank()) +
    guides(color = guide_legend(
      override.aes = list(linetype = "solid", size = 5)))

pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure1/env_ts.pdf", width = 9,height=7)
print(envs)
dev.off()


### sociodemographic time series  ###
## colored by province
cols <- colnames(sa_adm2_weekly)
sero_cols <- grep("Sero",cols, value=T)
age_cols <- c("age_lt6", "age_6t14", "age_15t64", "age_gt65")
type_cols <- c("other_count", "mening_count", "bact_count","WGS_PEN_SIR_Meningitis_R_count")
vaccine_cols <- c("pcv7", "pcv13", "nvt")
disease_cols <- c("disease")

soc_data <- sa_adm2_weekly %>%
  group_by(date) %>%
  summarise(across(
    all_of(c(disease_cols, age_cols, type_cols, vaccine_cols)),
    ~ sum(.x, na.rm = TRUE)
  ), .groups = "drop") %>%
  pivot_longer(
    cols = -c(date),
    names_to = "variable",
    values_to = "value"
  ) %>%
  mutate(
    facet_group = case_when(
      variable %in% disease_cols ~ "Disease",
      variable %in% age_cols ~ "Age Group",
      variable %in% type_cols ~ "Disease Type",
      variable %in% vaccine_cols ~ "Vaccine Type",
      TRUE ~ "Other"
    )
  )

yearly_summary <- soc_data %>%
  # mutate(year = year(as.Date(date))) %>%  # Convert IDate to Date
  group_by( variable, facet_group) %>%
  summarise(value = sum(value), .groups = "drop") %>%
  arrange(facet_group, variable, value)
# write.csv(yearly_summary, "/home/sbelman/Documents/env_sa_manuscript/figures/figure1/yearly_summary.csv", row.names = FALSE)

## factor for order
soc_data$facet_group <- factor(soc_data$facet_group, 
                               levels = c("Disease","Age Group", "Disease Type", "Vaccine Type"),
                               labels = c("Disease","Age\nGroup", "Disease\nType", "Vaccine\nType"))
soc_data$variable <- factor(soc_data$variable, 
                               levels = c(disease_cols, age_cols, type_cols, vaccine_cols))
## remove count label
soc_data$variable <- gsub("_count","",soc_data$variable)

 # color palettes
age_colors <- setNames(brewer.pal(4, "Set2"), age_cols)
disease_type_colors <- c("bact" = "#D7191C", "mening" = "#2C7BB6", "other" = "#1A9641","WGS_PEN_SIR_Meningitis_R"= "purple")
# disease_type_colors <- c("bact" = "#D7191C", "mening" = "#2C7BB6", "other" = "#1A9641")
pcv_colors <- c("nvt" = "#E75480", "pcv7" = "#522081", "pcv13" = "darkgreen")
disease_color <- c("disease" = "black")
sero_groups <- list(
  PCV7  = c("Sero14", "Sero19F", "Sero23F", "Sero4", "Sero6B", "Sero9V","Sero18C"),
  PCV13 = c("Sero1", "Sero19A", "Sero3", "Sero6A","Sero7F"),
  NVT   = c("Sero8", "Sero12F","Sero16F","Sero9N","Sero15A","Sero15BC")
)
## serotype colors ##
sero_colors <- unlist(lapply(names(sero_groups), function(g) {
  setNames(rep(pcv_colors[[tolower(g)]], length(sero_groups[[g]])), sero_groups[[g]])
}))
custom_colors <- c(disease_color, age_colors, disease_type_colors, pcv_colors, sero_colors)
# all_variables <- unique(soc_data$variable)
# missing_colors <- setdiff(all_variables, names(custom_colors))
# random_colors <- setNames(viridis(length(missing_colors), option = "D"), missing_colors)
final_color_map <- c(custom_colors)
soc_data$variable <- factor(soc_data$variable, 
                               levels = gsub("_count","",c(disease_cols, age_cols, type_cols, vaccine_cols, sero_cols)))
soc_data_2023 <- subset(soc_data,  soc_data$variable != "disease")
soc_ts_2023 <- ggplot(soc_data_2023) +
    geom_vline(xintercept=as.Date("2009-01-01"), linetype ="dashed",color = "black", alpha = 0.7)+
    geom_vline(xintercept=as.Date("2011-01-01"), linetype ="dashed",color = "black", alpha = 0.7)+
  geom_line(aes(x = date, y = value, group = interaction(variable), color = variable), alpha = 0.9) +
  facet_wrap(~ facet_group, scales = "free_y", ncol = 1, strip.position = "left")+
    scale_color_manual(values = final_color_map) +
  theme_bw() +
  ylab("case count") +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    strip.text.y = element_text(size = 14), axis.text = element_text(size=13)
  ) +
  guides(color = guide_legend(nrow = 11,override.aes = list(size = 4)))

pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure1/soc_ts_2023.pdf", width = 7.5,height=7)
print(soc_ts_2023)
dev.off()

### repeat for <2020
soc_data_2019 <- subset(soc_data, soc_data$date < as.Date("2020-01-01") & soc_data$variable != "disease")
socts <- ggplot(soc_data_2019) +
  geom_line(aes(x = date, y = value, group = interaction(variable), color = variable), alpha = 0.9) +
  facet_wrap(~ facet_group, scales = "free_y", ncol = 1, strip.position = "left")+
    scale_color_manual(values = final_color_map) +
  theme_bw() +
  ylab("case count") +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    strip.text.y = element_text(size = 14), axis.text = element_text(size=13)
  ) +
  guides(color = guide_legend(nrow = 10,override.aes = list(size = 4)))

pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure1/soc_ts_2019.pdf", width = 6,height=7)
print(socts)
dev.off()
```
## Figure 1D: disease separately with particulate matter and hurs
```{r}
sa_adm1_monthly<-fread(file="/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm1_monthly.csv")

  # Reshape data to long format
    df_provs <- sa_adm1_monthly %>%
      group_by(date) %>%
      summarise(
        pm2p5 = mean(pm2p5, na.rm = TRUE),
        hurs = mean(hurs, na.rm = TRUE),
        disease = sum(disease, na.rm = TRUE),
        tasmax = mean(tasmax, na.rm = TRUE)
      ) #%>%
      # mutate(disease_smooth = zoo::rollsum(disease, k = 7, fill = NA, align = "right"))      
    
    # clip extremes
    df_provs$pm2p5[which(df_provs$pm2p5 > 300)] <- 300
    # Scaling pm2p5 and hurs to the range of disease_smooth
    # max_disease <- max(df_provs$disease_smooth, na.rm = TRUE)
    max_disease <- max(df_provs$disease, na.rm = TRUE)


   ### Compute scale factor per location
    df_scaled <- df_provs %>%
      mutate(
        scale_factor_pm2p5 = max_disease / max(pm2p5, na.rm = TRUE),
        scale_factor_hurs = max_disease / max(hurs, na.rm = TRUE),
        scale_factor_tasmax = max_disease / max(tasmax, na.rm = TRUE),
        pm2p5_scaled = pm2p5 * scale_factor_pm2p5,
        hurs_scaled = hurs * scale_factor_hurs,
        tasmax_scaled = tasmax * scale_factor_tasmax
      )
    
    
# Plot with secondary axes using ggplot2
dis_env <- ggplot(df_scaled, aes(x = date)) +
  # geom_line(aes(y = disease_smooth, color = "Disease"), linewidth = 0.8) +
  geom_line(aes(y = disease, color = "Disease"), linewidth = 0.8) +
  geom_line(aes(y = pm2p5_scaled, color = "PM2.5"), alpha = 0.7, linetype = "solid",  linewidth = 0.6) +
  geom_line(aes(y = hurs_scaled, color = "Humidity"), alpha = 0.7, linetype = "solid" , linewidth = 0.6) +
  scale_color_manual(
    name = "Variable",
    values = c("Disease" = "black", "PM2.5" = "darkred", "Humidity" = "darkblue")
  ) +
  scale_y_continuous(
    name = "Disease case counts",
    sec.axis = sec_axis(~ . / df_scaled$scale_factor_pm2p5[1], name = "PM2.5 ug/m3"),
    # sec.axis = sec_axis(~ . / df_scaled$scale_factor_hurs[1], name = "Humidity")
  ) +
  theme_bw() +
  labs(x = "Date") +
  theme(
    axis.title.y.left = element_text(color = "black"),
    axis.title.y.right = element_text(color = "darkred"),
    legend.position = "top", legend.title = element_blank()
  )

# pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure1/disenv_ts_2023.pdf", width = 9,height=4)
# print(dis_env)
# dev.off()


dis_pm <- ggplot(df_scaled, aes(x = date)) +
      geom_vline(xintercept=as.Date("2009-01-01"), linetype ="dashed",color = "grey20", alpha = 0.7)+
    geom_vline(xintercept=as.Date("2011-01-01"), linetype ="dashed",color = "grey20", alpha = 0.7)+
  # geom_line(aes(y = disease_smooth, color = "Disease"), linewidth = 0.8) +
    geom_line(aes(y = disease, color = "Disease"), linewidth = 0.8, alpha = 0.7) +
  geom_line(aes(y = pm2p5_scaled, color = "PM2.5"), alpha = 0.7, linetype = "solid",  linewidth = 0.6) +
  scale_color_manual(
    name = "Variable",
    values = c("Disease" = "black", "PM2.5" = "darkred"), labels = c("Disease"="Disease","PM2.5"=expression("PM"["2.5"]))) +
  scale_y_continuous(
    name = "case counts",
    sec.axis = sec_axis(~ . / df_scaled$scale_factor_pm2p5[1], name = expression("PM"["2.5"]~mu*"g/"*m^3)),
    # sec.axis = sec_axis(~ . / df_scaled$scale_factor_hurs[1], name = "Humidity")
  ) +
  theme_bw() +
  labs(x = "Date") +
  theme(
    axis.title.y.left = element_text(color = "black"),
    axis.title.y.right = element_text(color = "darkred"),
    legend.position = "top", legend.title = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size=15), axis.title = element_text(size=15)
  )
dis_hurs <- ggplot(df_scaled, aes(x = date)) +
        geom_vline(xintercept=as.Date("2009-01-01"), linetype ="dashed",color = "grey20", alpha = 0.7)+
    geom_vline(xintercept=as.Date("2011-01-01"), linetype ="dashed",color = "grey20", alpha = 0.7)+
  # geom_line(aes(y = disease_smooth, color = "Disease"), linewidth = 0.8) +
  geom_line(aes(y = disease, color = "Disease"), linewidth = 0.8, alpha = 0.7) +
  geom_line(aes(y = hurs_scaled, color = "Humidity"), alpha = 0.7, linetype = "solid" , linewidth = 0.6) +
  scale_color_manual(
    name = "Variable",
    values = c("Disease" = "black",  "Humidity" = "darkblue")
  ) +
  scale_y_continuous(
    name = "case counts",
    # sec.axis = sec_axis(~ . / df_scaled$scale_factor_pm2p5[1], name = "PM2.5 ug/m3"),
    sec.axis = sec_axis(~ . / df_scaled$scale_factor_hurs[1], name = "Humidity")
  ) +
  theme_bw() +
  labs(x = "Date") +
  theme(
    axis.title.y.left = element_text(color = "black"),
    axis.title.y.right = element_text(color = "darkblue"),
    legend.position = "top", legend.title = element_blank(),
        axis.text.y = element_text(size=15), axis.title = element_text(size=15)
  )

pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure1/dispmhurs_ts_2023.pdf", width = 8,height=5)
print(dis_pm/dis_hurs)
dev.off()


#### replot this but just for those that are < 2020

df_scaled_2019 <- subset(df_scaled, df_scaled$date < as.Date("2020-01-01"))
dis_pm <- ggplot(df_scaled_2019, aes(x = date)) +
  # geom_line(aes(y = disease_smooth, color = "Disease"), linewidth = 0.8) +
    geom_line(aes(y = disease, color = "Disease"), linewidth = 0.8, alpha = 0.7) +
  geom_line(aes(y = pm2p5_scaled, color = "PM2.5"), alpha = 0.7, linetype = "solid",  linewidth = 0.6) +
  scale_color_manual(name = "Variable",values = c("Disease" = "black", "PM2.5" = "darkred"), labels = c("Disease"="Disease","PM2.5"=expression("PM"["2.5"]))) + scale_y_continuous(name = "case counts",sec.axis = sec_axis(~ . / df_scaled$scale_factor_pm2p5[1], name = expression("PM"["2.5"]~mu*"g/"*m^3))) +
  theme_bw() +
  labs(x = "Date") +
  theme(
    axis.title.y.left = element_text(color = "black"),
    axis.title.y.right = element_text(color = "darkred"),
    legend.position = "top", legend.title = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size=15), axis.title = element_text(size=15))

dis_hurs <- ggplot(df_scaled_2019, aes(x = date)) +
  # geom_line(aes(y = disease_smooth, color = "Disease"), linewidth = 0.8) +
  geom_line(aes(y = disease, color = "Disease"), linewidth = 0.8, alpha = 0.7) +
  geom_line(aes(y = hurs_scaled, color = "Humidity"), alpha = 0.7, linetype = "solid" , linewidth = 0.6) +
  scale_color_manual(name = "Variable",values = c("Disease" = "black",  "Humidity" = "darkblue")) +
  scale_y_continuous(name = "case counts",sec.axis = sec_axis(~ . / df_scaled$scale_factor_hurs[1], name = "Humidity")) +
  theme_bw() +
  labs(x = "Date") +
  theme(axis.title.y.left = element_text(color = "black"),axis.title.y.right = element_text(color = "darkblue"),legend.position = "top", legend.title = element_blank(),axis.text.y = element_text(size=15), axis.title = element_text(size=15))
pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure1/dispmhurs_ts_2019.pdf", width = 8,height=5)
print(dis_pm/dis_hurs)
dev.off()
```

Scripts for visualizing Main Figure 2
## Figures 2A-C
#### base model random effects
```{r}
time = "weekly"
space = "adm2"
endyear = 2019
data<-fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_",space,"_",time,"_lag_sc.csv"))
g <- inla.read.graph(filename = "/home/sbelman/Documents/env_sa_manuscript/input_datasets/shps/sa_adjacency_map.adj")
shp<-st_read("/home/sbelman/Documents/env_sa_manuscript/input_datasets/shps/gadm41_namematch_ZAF_2.shp")
basemod <- readRDS(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/base_models/base_model_",time,"_20092011_popdens_",space,"_",endyear,".rds"))
## spatial effects
N=52
mod<-basemod
  shp_a1<-shp
  if(is.null(shp_a1$areaid)){
    if(N==52){
    shp_a1$areaid <- shp_a1$GID_2
    }else{
      shp_a1$areaid <- shp_a1$GID_1
      
    }
  }
  shp_a1 <- shp_a1[order(shp_a1$areaid), ]
  cov<-mod$cov
  unstructured_effects <- mod$summary.random$id_u$mean[1:N]
  shp_a1$unstructured_effects<-unstructured_effects
  
  spat<-ggplot(data = shp_a1) +
    geom_sf(aes(fill = unstructured_effects),color=NA) +
    scale_fill_gradient2(low="blue",mid="white",high="red",midpoint = 0)+
    theme_classic() +
    labs(fill = "Random Effect")
## seasonal effects
  month_names <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", 
                         "October", "November", "December")
  mp <- basemod$summary.random$id_m
  mp$month_name<-month_names
  mp$month_name <- factor(mp$month_name, levels = c(month_names))
  colnames(mp)[grep("quant",colnames(mp))] <- c("lowerCI","median","upperCI")
  m <- ggplot(mp)+
        geom_line(aes(x = month_name, y = median, group=1))+
        geom_hline(yintercept=0, linetype="dashed", color="red")+
        geom_ribbon(aes(x = month_name, ymin = lowerCI, ymax = upperCI, group =1), alpha = 0.5)+
        theme_bw()+
        theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 11), axis.title = element_text(size = 11))+
        xlab("Month")+
        ylab("Seasonal Effect")
  
  yp <- basemod$summary.random$id_y
  yp$year <- 2004+ yp$ID
  year_n <- nrow(yp)/9 
  colnames(yp)[grep("quant",colnames(yp))] <- c("lowerCI","median","upperCI")
  reg_month_year <- rep(c("Eastern_Cape", "Free_State", "Gauteng",
                                                       "KwaZulu-Natal", "Limpopo", "Mpumalanga",
                                                       "North_West", "Northern_Cape", "Western_Cape"), each = year_n)
  yp$region <- factor(reg_month_year)
      y <- ggplot(yp)+
        geom_hline(yintercept=0, linetype="dashed", color="red")+
        geom_ribbon(aes(x = year, ymin = lowerCI, ymax = upperCI, group = region, ), alpha = 0.3)+
        geom_line(aes(x = year, y = median, group=region))+
        theme_bw()+
        theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 11), axis.title = element_text(size = 11))+
        xlab("Year")+
        ylab("Annual Effect")+
        labs(fill="Province",color = "Province")+
        facet_wrap(region ~.)

pdf(file=paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure2/res_weekly_adm2_",endyear,".pdf"), width = 12,height=4)
print(spat+m+y)
dev.off()
```
## Figure 2D: Fixed Effects
#### plot fixed effects from baseline model
```{r}
###### plot vaccination period and population_density
fixedeffs <- basemod$summary.fixed
colnames(fixedeffs) <- c("mean","sd","lowerCI","median","upperCI","mode","kld")
fixedeffs$covs <- c("int","postPCV7","postPCV13","population_density")

## adjust mean effect
fixedeffs$mean_all <- fixedeffs$mean
fixedeffs$mean_all[which(fixedeffs$covs=="postPCV7")] <- fixedeffs$mean[which(fixedeffs$covs=="postPCV7")] + fixedeffs$mean[which(fixedeffs$covs=="int")] 
fixedeffs$mean_all[which(fixedeffs$covs=="postPCV13")] <- fixedeffs$mean[which(fixedeffs$covs=="postPCV13")] + fixedeffs$mean[which(fixedeffs$covs=="int")] 

## adjust lower CI
fixedeffs$lowerCI_all <- fixedeffs$lowerCI
fixedeffs$lowerCI_all[which(fixedeffs$covs=="postPCV7")] <- fixedeffs$lowerCI[which(fixedeffs$covs=="postPCV7")] + fixedeffs$lowerCI[which(fixedeffs$covs=="int")] 
fixedeffs$lowerCI_all[which(fixedeffs$covs=="postPCV13")] <- fixedeffs$lowerCI[which(fixedeffs$covs=="postPCV13")] + fixedeffs$lowerCI[which(fixedeffs$covs=="int")] 

## adjust upper CI
fixedeffs$upperCI_all <- fixedeffs$upperCI
fixedeffs$upperCI_all[which(fixedeffs$covs=="postPCV7")] <- fixedeffs$upperCI[which(fixedeffs$covs=="postPCV7")] + fixedeffs$upperCI[which(fixedeffs$covs=="int")] 
fixedeffs$upperCI_all[which(fixedeffs$covs=="postPCV13")] <- fixedeffs$upperCI[which(fixedeffs$covs=="postPCV13")] + fixedeffs$upperCI[which(fixedeffs$covs=="int")] 

## change to factors
fixedeffs$covs <- factor(fixedeffs$covs, levels = c("int","postPCV7","postPCV13","population_density"),
                         labels = c("prePCV","postPCV7","postPCV13","population density"))

fixedeffs$OR <- exp(fixedeffs$mean_all)
fixedeffs$OR_lower <- exp(fixedeffs$lowerCI_all)
fixedeffs$OR_upper <- exp(fixedeffs$upperCI_all)
fixedeffs$type <- ifelse(fixedeffs$covs=="population density","population","vaccine")
## plot these effects for different periods
vax_colors <- c("prePCV" = "steelblue", "postPCV7"= "#FFC000", "postPCV13" = "darkorange", "population density" = "grey20")
fixedplot <- ggplot(fixedeffs) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red",alpha = 0.5) + 
  geom_pointrange(aes(x=covs, y = mean_all, ymin = lowerCI_all, ymax = upperCI_all, group = type, color = covs))+
  theme_bw()+
  ylim(-3.5,0)+
  scale_color_manual(values = vax_colors) +
  xlab("Covariate") +
  ylab("Fixed Effect") +
  theme(legend.title = element_blank(), axis.text = element_text(size =11), axis.title = element_text(size=13),
        axis.text.x = element_text(angle=45, hjust =1))

scaleFUN <- function(x) sprintf("%.4f", x)
popinset <- ggplot(fixedeffs[which(fixedeffs$type=="population"),]) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red",alpha = 0.5) + 
  geom_pointrange(aes(x=covs, y = mean_all, ymin = lowerCI_all, ymax = upperCI_all, group = type, color = covs))+
  theme_classic()+
  # ylim(-3.5,0)+
  scale_color_manual(values = vax_colors) +
  scale_y_continuous(labels=scaleFUN) +
  xlab("") +
  ylab("Fixed Effect") +
  theme(legend.title = element_blank(), axis.text = element_text(size =10), axis.title = element_text(size=10),
        axis.text.x = element_blank(), legend.position = "none")
pdf(file=paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure2/population_densityinset_",endyear,"_fixedeffects.pdf"), width = 1.5,height=1.5)
print(popinset)
dev.off()

pdf(file=paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure2/res_weekly_adm2_",endyear,"_fixedeffects.pdf"), width = 4,height=4)
print(fixedplot)
dev.off()
```

## Figure 3
Scripts for visualizing Main Figure 3 and associated supplementary figures
## LOAD DATA FOR AIR POLLUTION AND METEOROLOGICAL MODELS
```{r}
# ##  VISUALIZE BIVARIABLE OUTPUTS 
st_vec <- c("weekly_adm1","weekly_adm2")
# st_vec <- c("weekly_adm2")
endyear = 2019

### read in bivariable
stlist <- list()
for(s in 1:length(st_vec)){
  m1 <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/nointeraction_results_fits_",st_vec[s],"_bivariable_8_2019_templag0humlag3.csv"))
  m1$space_time <- st_vec[s]
  stlist[[s]] <- m1
}
fit_list_biv <- rbindlist(stlist)
fit_list_biv$cov <- gsub("_lag0","",fit_list_biv$covariate)
# fit_list_biv$cov2 <- gsub("_lag3","",fit_list_biv$cov2)

### read in univariable
stlist <- list()
for(s in 1:length(st_vec)){
  m1 <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/univariable/nointeraction_results_fits_",st_vec[s],"_8_2019.csv"))
  m1$space_time <- st_vec[s]
  stlist[[s]] <- m1
}
fit_list <- rbindlist(stlist)
fit_list$cov <- gsub("_lag0","",fit_list$covariate)
fit_list$cov2 <- "none"
fit_list <- rbind(fit_list, fit_list_biv)
fit_list$lag_week <- paste0("Week",fit_list$lag_num)
fit_list$lag_week <- factor(fit_list$lag_week, levels = paste0("Week",seq(0,12,1)), labels = paste0("Week",seq(0,12,1)))
# convert to relative risks
fit_list$rr <- exp(fit_list$fit)
fit_list$lowerCI_rr <- exp(fit_list$lowerCI)
fit_list$upperCI_rr <- exp(fit_list$upperCI)
fit_list$cumrr <- exp(fit_list$cumulative_fit)
fit_list$cumlowerCI_rr <- exp(fit_list$cum_lowerCI)
fit_list$cumupperCI_rr <- exp(fit_list$cum_upperCI)

# add in what these actually are
fit_list$cov2[which(fit_list$cov2 == "absh_lag3")] <- "absh_lag3 + tas_lag0"
fit_list$cov2[which(fit_list$cov2 == "hurs_lag3")] <- "hurs_lag3 + tas_lag0"

# set labels 
color_labels <- c("absh_lag3 + tas_lag0"="lightblue", "hurs_lag3 + tas_lag0" = "darkblue","tasmax_lag0" = "brown","tas_lag0" = "orange", "none" = "black")
linetype_labels <- c("absh_lag3 + tas_lag0"="solid", "hurs_lag3 + tas_lag0" = "solid","tasmax_lag0" = "solid","tas_lag0" = "solid", "none" = "solid")

# 
# color_labels <- c("pm2p5" = "red", "pm10" = "orange","so2"= "pink","absh"="lightblue", "hurs" = "darkblue","tasmax" = "purple","tasmin" = "brown","tas" = "orange", "none" = "darkgreen")
# linetype_labels <-c("pm2p5" = "solid", "pm10" = "solid","so2" = "solid","absh"="solid", "hurs" = "solid","tasmax" = "solid", "tas" = "solid", "tasmin" = "solid", "none" = "longdash")
```

## FIGURE 3:  Cumulative Plot for each
```{r}
space_time_str <- "weekly_adm2"
data_unscaled <- fread("/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm2_weekly_lag.csv")
data_unscaled_aggregate <- data_unscaled %>%
  group_by(district) %>%
  summarise(pm2p5 = mean(pm2p5_lag0,na.rm=T),
            pm10 = mean(pm10_lag0,na.rm=T),
            hurs = mean(hurs_lag0,na.rm=T),
            absh = mean(absh_lag0,na.rm=T),
            tas = mean(tas_lag0,na.rm=T),
            so2 = mean(so2_lag0, na.rm=T))
# TEMPERATURE
tmp <- subset(fit_list, fit_list$cov%in%c("tas","tasmax","tasmin") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
# tmp$cov2 <- factor(tmp$cov2, levels = c("pm2p5","pm10","so2","tasmax","none"))
plot_temp_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov), color=cov),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov), fill=cov),alpha=0.3)+
    geom_rug(data= data_unscaled, aes(x = tas_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  scale_color_manual(values = c("tasmin" = "lightblue","tas" = "#BA8E23", "tasmax"="darkred"))+
  scale_fill_manual(values = c("tasmin" = "lightblue","tas" = "#BA8E23", "tasmax"="darkred"))+
  # scale_linetype_manual(values = linetype_labels)+
  labs(color = "Temperature Metric", fill = "Temperature Metric")+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10", limits = c(0.4,1.1), breaks = c(0.4, 1, 1.1))+
    scale_y_continuous(trans="log10", limits = c(0.7,1.2), breaks = c(0.8, 1, 1.2))+
  ggtitle("Temperature")+
  xlab("°C")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "right")

# HURS
tmp <- subset(fit_list, fit_list$cov%in%c("hurs") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
# tmp$cov2 <- factor(tmp$cov2, levels = c("pm2p5","pm10","so2","tasmax","none"))
plot_hurs_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov2), color=cov2, linetype = cov2),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov2), fill=cov2),alpha=0.3)+
    geom_rug(data= data_unscaled, aes(x = hurs_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  scale_color_manual(values = color_labels)+
  scale_fill_manual(values = color_labels)+
  scale_linetype_manual(values = linetype_labels)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.6,1.35), breaks = c(0.6, 1, 1.3))+
  ggtitle("Relative Humidity")+
  labs(linetype = "additive variable", color = "additive variable", fill = "additive variable")+
  # facet_wrap(.~lag_week, nrow = 1)+
  xlab("Percent")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "none")
# ABSH
tmp <- subset(fit_list, fit_list$cov%in%c("absh") &  fit_list$lag_num %in% c(3) & fit_list$space_time%in%space_time_str)
# tmp$cov2 <- factor(tmp$cov2, levels = c("pm2p5","pm10","so2","tasmax","none"))
plot_absh_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov2), color=cov2, linetype = cov2),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov2), fill=cov2),alpha=0.3)+
  geom_rug(data= data_unscaled, aes(x = absh_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  scale_color_manual(values = color_labels)+
  scale_fill_manual(values = color_labels)+
  scale_linetype_manual(values = linetype_labels)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.6,1.35), breaks = c(0.6, 1, 1.3))+
  ggtitle("Absolute Humidity")+
  labs(linetype = "additive variable", color = "additive variable", fill = "additive variable")+
  # facet_wrap(.~lag_week, nrow = 1)+
  xlab(expression(paste('Water (g/m'^3,')' )))+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "none")
# PM2.5
tmp <- subset(fit_list, fit_list$cov%in%c("pm2p5") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
tmp$cov2 <- factor(tmp$cov2, levels = c("absh_lag3 + tas_lag0","hurs_lag3 + tas_lag0","tas_lag0","tasmax_lag0","none"))
plot_pm2p5_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov2), color=cov2, linetype = cov2),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov2), fill=cov2),alpha=0.04)+
        geom_rug(data= data_unscaled, aes(x = pm2p5_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  scale_color_manual(values = color_labels)+
  scale_fill_manual(values = color_labels)+
  scale_linetype_manual(values = linetype_labels)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.8,1.9), breaks = c(0.8, 1, 1.8))+
  ggtitle("PM2.5")+
  labs(linetype = "additive variable", color = "additive variable", fill = "additive variable")+
  # facet_wrap(.~lag_week, nrow = 1)+
  xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "right")
# PM10
tmp <- subset(fit_list, fit_list$cov%in%c("pm10") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
tmp$cov2 <- factor(tmp$cov2, levels = c("absh_lag3 + tas_lag0","hurs_lag3 + tas_lag0","tas_lag0","tasmax_lag0","none"))
plot_pm10_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov2), color=cov2, linetype = cov2),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov2), fill=cov2),alpha=0.04)+
      geom_rug(data= data_unscaled, aes(x = pm10_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  scale_color_manual(values = color_labels)+
  scale_fill_manual(values = color_labels)+
  scale_linetype_manual(values = linetype_labels)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.8,1.9), breaks = c(0.8, 1, 1.8))+
  ggtitle("PM10")+
  labs(linetype = "additive variable", color = "additive variable", fill = "additive variable")+
  # facet_wrap(.~lag_week, nrow = 1)+
  xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "none")

# Sulfur Dioxide
tmp <- subset(fit_list, fit_list$cov%in%c("so2") &  fit_list$lag_num %in% c(3) & fit_list$space_time%in%space_time_str)
tmp$cov2 <- factor(tmp$cov2, levels = c("absh_lag3 + tas_lag0","hurs_lag3 + tas_lag0","tas_lag0","tasmax_lag0","none"))
plot_so2_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov2), color=cov2, linetype = cov2),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov2), fill=cov2),alpha=0.04)+
  geom_rug(data= data_unscaled, aes(x = so2_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  scale_color_manual(values = color_labels)+
  scale_fill_manual(values = color_labels)+
  scale_linetype_manual(values = linetype_labels)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.8,1.9), breaks = c(0.8, 1, 1.8))+
  ggtitle("Sulfur Dioxide")+
  labs(linetype = "additive variable", color = "additive variable", fill = "additive variable")+
  # facet_wrap(.~lag_week, nrow = 1)+
  xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "none")


### save each individually
pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/pm2p5_cumulative_additive_sa_",space_time_str,".pdf"),width  =3.5,height=3.5)
print(plot_pm2p5_cum)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/so2_cumulative_additive_sa_",space_time_str,".pdf"),width  =3.5,height=3.5)
print(plot_so2_cum)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/pm10_cumulative_additive_sa_",space_time_str,".pdf"),width  =3.5,height=3.5)
print(plot_pm10_cum)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/absh_cumulative_additive_sa_",space_time_str,".pdf"),width  =3.5,height=3.5)
print(plot_absh_cum)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/hurs_cumulative_additive_sa_",space_time_str,".pdf"),width  =3.5,height=3.5)
print(plot_hurs_cum)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/tas_cumulative_additive_sa_",space_time_str,".pdf"),width  =3.5,height=3.5)
print(plot_temp_cum)
dev.off()

### save together
pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/cumulative_additive_sa_",space_time_str,".pdf"),width  =10,height=7)
(plot_pm2p5_cum + plot_pm10_cum + plot_so2_cum)/
  (plot_hurs_cum + plot_absh_cum + plot_temp_cum)
dev.off()
# 
# if(space_time_str=="weekly_adm2"){
# (plot_pm2p5_cum + plot_pm10_cum)/
#   (plot_hurs_cum + plot_absh_cum)
#   ggsave(file = paste0("/home/sbelman/Documents/env_sa_manuscript/figures/cumulative_additive_sa_",space_time_str,".png"), width = 9, height = 7)
#   }else{
#    (plot_pm2p5_cum + plot_pm10_cum + plot_hurs_cum )
#     ggsave(file = paste0("/home/sbelman/Documents/env_sa_manuscript/figures/cumulative_additive_sa_",space_time_str,".png"), width = 13, height = 4)
# 
#   }

# # Maximum temperature
# space_time_str <- "weekly_adm2"
# tmptemp <- subset(fit_list, fit_list$cov%in%c("tasmin","tasmax") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
# plot_temp_cum <- ggplot(tmptemp)+
#   geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov), color=cov),alpha=0.5)+
#   geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
#   geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov), fill=cov),alpha=0.04)+
#   theme_bw()+
#   xlab("Var")+
#   ylab("Relative Risk")+
#   scale_y_continuous(trans="log10")+
#   scale_color_manual(values = c("tasmax"="darkred","tasmin"="blue"), labels = c("tasmax"="Max.Temp","tasmin"="Min.Temp")) + 
#     scale_fill_manual(values = c("tasmax"="darkred","tasmin"="blue"), labels = c("tasmax"="Max.Temp","tasmin"="Min.Temp")) + 
# 
#   # scale_y_continuous(trans="log10", limits = c(0.8,1.9), breaks = c(0.8, 1, 1.8))+
#   ggtitle("Temperature")+
#   # facet_wrap(.~lag_week, nrow = 1)+
#   xlab("Celsius")+
#   theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
#         strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
#         legend.position = "right",legend.title = element_blank())
# plot_temp_cum
# ggsave("/home/sbelman/Documents/env_sa_manuscript/figures/cumulative_temperature_maxmin_weekly_adm2.png", width = 7, height = 5)

# TEMPERATURE
tmp <- subset(fit_list, fit_list$cov%in%c("tasmin") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
# tmp$cov2 <- factor(tmp$cov2, levels = c("pm2p5","pm10","so2","tasmax","none"))
plot_tempmin_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov), color=cov),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov), fill=cov),alpha=0.3)+
    geom_rug(data= data_unscaled, aes(x = tas_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  scale_color_manual(values = c("tasmin" = "#4169E1"))+
  scale_fill_manual(values = c("tasmin" = "#4169E1"))+
  # scale_linetype_manual(values = linetype_labels)+
  labs(color = "Temperature Metric", fill = "Temperature Metric")+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10", limits = c(0.8,1.09), breaks = c(0.8, 1, 1.1))+
    # scale_y_continuous(trans="log10", limits = c(0.7,1.1), breaks = c(0.8, 1, 1.1))+
  ggtitle("Temperature")+
  xlim(-4,15)+
  xlab("°C")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "none")
plot_tempmin_cum
```



### FIGURE S5
### Plot with lags across 8 weeks
```{r}
# Temperature
tmptas <- subset(fit_list, fit_list$cov%in%c("tasmin","tas","tasmax") & fit_list$lag_num %in% c(0:8) & fit_list$space_time=="weekly_adm2")

# tmphurs$cov2 <- factor(tmphurs$cov2, levels = c("pm2p5","pm10","so2","absh","tasmax","none"))
plot_tas <- ggplot(tmptas)+
  geom_line(aes(x=var,y=rr,group=interaction(lag_week,space_time, cov), color=cov),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(lag_week,space_time,cov), fill=cov),alpha=0.3)+
  scale_color_manual(values = c("tasmin" = "lightblue","tas" = "#BA8E23", "tasmax"="darkred"))+
  scale_fill_manual(values = c("tasmin" = "lightblue","tas" = "#BA8E23", "tasmax"="darkred"))+
  # scale_linetype_manual(values = linetype_labels)+
  labs(color = "Temperature Metric", fill = "Temperature Metric")+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10", limits = c(0.7,1.2), breaks = c(0.8, 1, 1.2))+
  ggtitle("Temperature")+
  facet_wrap(.~lag_week, nrow = 1)+
  xlab("°C")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13), legend.position = "right")


# Relative humidity
tmphurs <- subset(fit_list, fit_list$cov%in%c("hurs") & fit_list$lag_num %in% c(0:8) & fit_list$space_time=="weekly_adm2")

# tmphurs$cov2 <- factor(tmphurs$cov2, levels = c("pm2p5","pm10","so2","absh","tasmax","none"))
plot_hurs <- ggplot(tmphurs)+
  geom_line(aes(x=var,y=rr,group=interaction(lag_week,space_time, cov2), color=cov2, linetype = cov2),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(lag_week,space_time,cov2), fill=cov2),alpha=0.3)+
  scale_color_manual(values = color_labels)+
  scale_fill_manual(values = color_labels)+
  scale_linetype_manual(values = linetype_labels)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.85,1.15), breaks = c(0.85, 1, 1.15))+
  ggtitle("Relative Humidity")+
  labs(linetype = "additive variable", color = "additive variable", fill = "additive variable")+
  facet_wrap(.~lag_week, nrow = 1)+
  xlab("Percent")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13), legend.position = "right")


# ABSH
tmpabsh <- subset(fit_list, fit_list$cov%in%c("absh") &  fit_list$lag_num %in% c(0:12) & fit_list$space_time=="weekly_adm2")
# tmpabsh$cov2 <- factor(tmpabsh$cov2, levels = c("pm2p5","pm10","so2","hurs","tasmax","none"))
plot_absh <- ggplot(tmpabsh)+
  geom_line(aes(x=var,y=rr,group=interaction(lag_week,space_time, cov2), color=cov2, linetype = cov2),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(lag_week,space_time,cov2), fill=cov2),alpha=0.3)+
  scale_color_manual(values = color_labels)+
  scale_fill_manual(values = color_labels)+
  scale_linetype_manual(values = linetype_labels)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.85,1.15), breaks = c(0.85, 1, 1.15))+
  ggtitle("Absolute Humidity")+
  labs(linetype = "additive variable", color = "additive variable", fill = "additive variable")+
  facet_wrap(.~lag_week, nrow = 1)+
  xlab(expression(paste('Water (g/m'^3,')' )))+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13), legend.position = "right")


### pm2.5
tmppm2 <- subset(fit_list, fit_list$cov%in%c("pm2p5") & fit_list$lag_num %in% c(0:12) & fit_list$space_time=="weekly_adm2")
tmppm2$cov2 <- factor(tmppm2$cov2, levels = c("absh_lag3 + tas_lag0","hurs_lag3 + tas_lag0","tas_lag0","tasmax_lag0","none"))
plot_pm2p5 <- ggplot(tmppm2)+
  geom_line(aes(x=var,y=rr,group=interaction(lag_week,space_time, cov2), color=cov2, linetype = cov2),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(lag_week,space_time,cov2), fill=cov2),alpha=0.05)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.95,1.10), breaks = c(0.95, 1, 1.10))+
  scale_color_manual(values = color_labels)+
  scale_fill_manual(values = color_labels)+
  scale_linetype_manual(values = linetype_labels)+
  ggtitle("PM 2.5")+
  labs(linetype = "additive variable", color = "additive variable", fill = "additive variable")+
  facet_wrap(.~lag_week, nrow = 1)+
  xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13))


### PM10
tmppm1 <- subset(fit_list, fit_list$cov%in%c("pm10") & fit_list$lag_num %in% c(0:12) & fit_list$space_time=="weekly_adm2")
tmppm1$cov2 <- factor(tmppm1$cov2, levels = c("absh_lag3 + tas_lag0","hurs_lag3 + tas_lag0","tas_lag0","tasmax_lag0","none"))
plot_pm10 <- ggplot(tmppm1)+
  geom_line(aes(x=var,y=rr,group=interaction(lag_week,space_time, cov2), color=cov2, linetype=cov2),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(lag_week,space_time,cov2), fill=cov2),alpha=0.05)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.85,1.15), breaks = c(0.85, 1, 1.15))+
  scale_color_manual(values = color_labels)+
  scale_fill_manual(values = color_labels)+
  scale_linetype_manual(values = linetype_labels)+
  ggtitle("PM 10")+
  labs(linetype = "additive variable", color = "additive variable", fill = "additive variable")+
  facet_wrap(.~lag_week, nrow = 1)+
  xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13))

# SO2
tmpso2 <- subset(fit_list, fit_list$cov%in%c("so2")  & fit_list$lag_num %in% c(0:12) & fit_list$space_time=="weekly_adm2")
tmpso2$cov2 <- factor(tmpso2$cov2, levels = c("absh_lag3 + tas_lag0","hurs_lag3 + tas_lag0","tas_lag0","tasmax_lag0","none"))

plot_so2 <- ggplot(tmpso2)+
  geom_line(aes(x=var,y=rr,group=interaction(lag_week,space_time, cov2), color=cov2, linetype = cov2),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(lag_week,space_time,cov2), fill=cov2),alpha=0.04)+
  scale_color_manual(values = color_labels)+
  scale_fill_manual(values = color_labels)+
  scale_linetype_manual(values = linetype_labels)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.85,1.15), breaks = c(0.85, 1, 1.15))+
  ggtitle("Sulfur Dioxide")+
  labs(linetype = "additive variable", color = "additive variable", fill = "additive variable")+
  facet_wrap(.~lag_week, nrow = 1)+
  xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13))
# plot_so2
# plot_pm2p5/plot_pm10
# plot_hurs/plot_absh
space_time_str <- "weekly_adm2"
pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/pm2p5pm10_weeklags_additive_sa_",space_time_str,".pdf"),width = 10.5,height=4)
(plot_pm2p5/ plot_pm10)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/abshhurs_weeklags_additive_sa_",space_time_str,".pdf"),width = 10,height=4)
(plot_hurs/plot_absh)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/so2_weeklags_additive_sa_",space_time_str,".pdf"),width = 10.5,height=2)
(plot_so2)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/tas_weeklags_additive_sa_",space_time_str,".pdf"),width = 10,height=2)
(plot_tas)
dev.off()
```

## Scripts for visualizing Main Figure 4
#### SUBSET BY MULTIPLE OUTCOME MODELS
```{r}
## multiple outcomes
    dlnm_multiout_demog <- fread("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/suboutcomes/dlnm_subsetoutcomes_results_fits_weekly_adm2_demographic_2019.csv")
    dlnm_multiout_sero <- fread("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/suboutcomes/dlnm_subsetoutcomes_results_fits_weekly_adm2_serotype_2019.csv")
    dlnm_multiout_sero$cov <- gsub("_lag0","",dlnm_multiout_sero$covariate)
    dlnm_multiout <- rbind(dlnm_multiout_demog, dlnm_multiout_sero)
    dlnm_multiout$cov <- gsub("_lag0","",dlnm_multiout$covariate)
    max_lag = 8
    ## no interaction single outcome of disease
    dlnm_norm <- fread("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/nointeraction_results_fits_weekly_adm2_bivariable_8_2019_templag0humlag3.csv")
    dlnm_norm <- subset(dlnm_norm, dlnm_norm$cov2=="absh_lag3") ## subset by absh_lag3 and temperature at lag0
    dlnm_norm$cov <- gsub("_lag0","",dlnm_norm$covariate)
    dlnm_norm_fit <- dlnm_norm[,c("fit","lowerCI","upperCI","cumulative_fit","cum_lowerCI","cum_upperCI","cov","predvar","lag_num")]
    colnames(dlnm_norm_fit) <- c("fit_dis","lowerCI_dis","upperCI_dis","cumulative_fit_dis","cum_lowerCI_dis","cum_upperCI_dis", "cov","predvar","lag_num")

    ## pcv, nvt outcomes
    ## pcv7
    dlnm_pcv7 <- subset(dlnm_multiout, dlnm_multiout$GPSC%in%c("pcv7"))
    dlnm_pcv7 <- dlnm_pcv7[,c("fit","lowerCI","upperCI","cov","predvar","lag_num")]
    colnames(dlnm_pcv7) <- c("fit_pcv7","lowerCI_pcv7","upperCI_pcv7", "cov","predvar","lag_num")
    ##pcv13
    dlnm_pcv13 <- subset(dlnm_multiout, dlnm_multiout$GPSC%in%c("pcv13"))
    dlnm_pcv13 <- dlnm_pcv13[,c("fit","lowerCI","upperCI","cov","predvar","lag_num")]
    colnames(dlnm_pcv13) <- c("fit_pcv13","lowerCI_pcv13","upperCI_pcv13", "cov","predvar","lag_num")
    ## nvt
    dlnm_nvt <- subset(dlnm_multiout, dlnm_multiout$GPSC%in%c("nvt"))
    dlnm_nvt <- dlnm_nvt[,c("fit","lowerCI","upperCI","cov","predvar","lag_num")]
    colnames(dlnm_nvt) <- c("fit_nvt","lowerCI_nvt","upperCI_nvt", "cov","predvar","lag_num")
    dlnm_vt <- left_join(dlnm_pcv7,dlnm_pcv13, by = c("cov","predvar","lag_num"))
    dlnm_vt <- left_join(dlnm_vt,dlnm_nvt, by = c("cov","predvar","lag_num"))
    dlnm_multiout <- left_join(dlnm_multiout,dlnm_vt, by = c("cov","predvar","lag_num"))

    ## merge no interaction and single outcome here
    dlnm_multiout <- left_join(dlnm_multiout,dlnm_norm_fit, by = c("cov","predvar","lag_num"))
    fits <- subset(dlnm_multiout, dlnm_multiout$cov=="pm2p5" )
    ## clean up
    fits$lag_week <- paste0("Week",fits$lag_num)
    fits$lag_week <- factor(fits$lag_week, levels = paste0("Week",seq(0,max_lag,1)), labels = paste0("Week",seq(0,max_lag,1)))
    fits$rr <- exp(fits$fit)
    fits$lowerCI_rr <- exp(fits$lowerCI)
    fits$upperCI_rr <- exp(fits$upperCI)

    ## clean up for all covs
    fits_covs <- dlnm_multiout
    fits_covs$lag_week <- paste0("Week",fits_covs$lag_num)
    fits_covs$lag_week <- factor(fits_covs$lag_week, levels = paste0("Week",seq(0,max_lag,1)), labels = paste0("Week",seq(0,max_lag,1)))
    fits_covs$rr <- exp(fits_covs$fit)
    fits_covs$lowerCI_rr <- exp(fits_covs$lowerCI)
    fits_covs$upperCI_rr <- exp(fits_covs$upperCI)
    fits_covs$cum_rr <- exp(fits_covs$cumulative_fit)
    fits_covs$cum_lowerCI_rr <- exp(fits_covs$cum_lowerCI)
    fits_covs$cum_upperCI_rr <- exp(fits_covs$cum_upperCI)
    
    ### serotype
    groups <- unique(fits$GPSC)
    outcomes <- grep("Sero",groups, value = TRUE)
    fits_sero <- subset(fits, fits$GPSC%in%outcomes & fits$lag_num%in%seq(0,max_lag,1))
    fits_pcv_type <- subset(fits, fits$GPSC%in%c("pcv7","pcv13","nvt")& fits$lag_num%in%seq(0,max_lag,1))
    fits_distypes <- subset(fits, fits$GPSC%in%c("bact_count","mening_count","other_count","WGS_PEN_SIR_Meningitis_R_count")& fits$lag_num%in%seq(0,max_lag,1))
    fits_age <- subset(fits, fits$GPSC%in%c("age_gt65","age_lt6","age_6t14","age_15t64","female_count")& fits$lag_num%in%seq(0,max_lag,1))
    
    ## label group
    pcv13_vec <- paste0("Sero",c( "1", "3", "5", "6A", "7F", "19A"),"_count")
    pcv7_vec <- paste0("Sero",c("4", "6B", "9V", "14", "18C", "19F", "23F"),"_count")
    nvt_vec <- paste0("Sero",c("12F","15A","15BC","16F","8","9N"),"_count")
    fits_sero$pcv_type <- ifelse(fits_sero$GPSC%in%pcv7_vec,"PCV7",
                            ifelse(fits_sero$GPSC%in%pcv13_vec,"PCV13","NVT"))

    fits_sero$pcv_type <- factor(fits_sero$pcv_type, levels = c("PCV7","PCV13","NVT"))
    fits_sero$GPSC <- gsub("_count","",fits_sero$GPSC)
```

#### Figure 4 and Figure S15 age distributions
```{r}
  fits_age_cum <- fits_age[which(fits_age$lag_num==3),]
  fits_age_cum$GPSC <- factor(fits_age_cum$GPSC, levels = c("age_lt6","age_6t14","age_15t64","age_gt65", "age_gt80","female_count"))
  

  plot_age <- ggplot(fits_age_cum)+
          geom_line(aes(x=predvar,y=exp(cumulative_fit_dis),group=interaction(lag_week,GPSC,cov)), color="black", alpha=0.7, linetype = "dashed")+
          geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
      geom_line(aes(x=predvar,y=exp(cumulative_fit),group=interaction(lag_week,GPSC,cov), color=GPSC),alpha=0.7)+
      geom_ribbon(aes(x=predvar, ymin=exp(cum_lowerCI),ymax=exp(cum_upperCI), group=interaction(lag_week,GPSC,cov), fill = GPSC),alpha=0.2)+
      # geom_ribbon(aes(x=predvar, ymin=exp(lowerCI_dis),ymax=exp(upperCI_dis), group=interaction(lag_week,GPSC)), fill = "black" ,alpha=0.1)+
      theme_bw()+
      xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
      ylab("Relative Risk")+
      # scale_y_continuous(trans="log10", limits = c(0.9,1.25), breaks = c(0.9, 1, 1.25))+
      scale_y_continuous(trans="log10")+
      ggtitle("PM2.5")+
      # facet_wrap(GPSC~lag_week , ncol=13)+
      labs(color="Age Group",fill="Age Group")+
      theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
            strip.text = element_text(size=9), axis.text.x = element_text(angle = 45,hjust=1, size=13), 
            legend.position = "right")
    plot_age
    ggsave(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/age_pm2p5_cumulative.png", width = 5,height=3)
    
pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/age_pm2p5_cumulative.pdf", width = 5,height=3)
print(plot_age)
dev.off()

#### across lags
     fits_age_lag <- fits_age[which(fits_age$lag_num<9),]
     fits_age_lag$GPSC <- factor(fits_age_lag$GPSC, levels = c("age_lt6","age_6t14","age_15t64","age_gt65", "age_gt80","female_count"))
     
     ## limit to plot
     fits_age_lag$upperCI[((fits_age_lag$upperCI>0.18))]<- 0.18

   plot_agelag <- ggplot(fits_age_lag)+
      geom_line(aes(x=predvar,y=exp(fit),group=interaction(lag_week,GPSC,cov), color=GPSC),alpha=0.7)+
      geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
      geom_ribbon(aes(x=predvar, ymin=exp(lowerCI),ymax=exp(upperCI), group=interaction(lag_week,GPSC,cov), fill = GPSC),alpha=0.1)+
      # geom_line(aes(x=predvar,y=exp(fit_dis),group=interaction(lag_week,GPSC)), color="black", alpha=0.7)+
      # geom_ribbon(aes(x=predvar, ymin=exp(lowerCI_dis),ymax=exp(upperCI_dis), group=interaction(lag_week,GPSC)), fill = "black" ,alpha=0.1)+
      theme_bw()+
      xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
      ylab("Relative Risk")+
      scale_y_continuous(trans="log10", limits = c(0.85,1.3))+
      #     scale_color_manual(labels = c("bact_count"="bacteremia","mening_count"="meningitis","other_count"="other invasive disease"), 
      #                    values = c("bact_count"="#D7191C","mening_count"="#2C7BB6","other_count"="#1A9641"))+
      # scale_fill_manual(labels = c("bact_count"="bacteremia","mening_count"="meningitis","other_count"="other invasive disease"),
      #                   values =  c("bact_count"="#D7191C","mening_count"="#2C7BB6","other_count"="#1A9641"))+
      ggtitle("PM2.5")+
      labs(color="Disease Type",fill="Disease Type")+
      theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
            strip.text = element_text(size=9), axis.text.x = element_text(angle = 45,hjust=1, size=13), 
            legend.position = "right")+
           facet_grid(GPSC~lag_num, scales = "free_y")
   
      pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/age_pm2p5_weekly.pdf", width = 12,height=6)
      print(plot_agelag)
      dev.off()

```
## Figure 4A
### Figure 4
#### plot the relative risk ratio for each sub model as compared to the main models with the cumulative effect
```{r}
# Filter to just age groups (GPSC column) — adapt if needed
fits_cum_pm2p5 <- fits[which(fits$var>45&fits$var<55),]
fits_cum_pm2p5 <- fits[which(fits$var==50),]

age_rr <- fits_cum_pm2p5 %>%
  # filter(GPSC %in% c("age_lt5", "age_5t18", "age_18t64", "age_gt65","female_count","mening_count","bact_count","other_count")) %>%
  mutate(
    RR = exp(cumulative_fit),
    RR_lower = exp(cum_lowerCI),
    RR_upper = exp(cum_upperCI)
  )
# Get the reference RR (age_lt5)
ref_rr <- age_rr %>%
  filter(GPSC == "age_lt6") %>%
  pull(RR)

### using the disease value
ref_cumfit_dis = (unique(fits_cum_pm2p5$cumulative_fit_dis))
ref_cumfit_dislower = (unique(fits_cum_pm2p5$cum_lowerCI_dis))
ref_cumfit_disupper = (unique(fits_cum_pm2p5$cum_upperCI_dis))

# Calculate RR ratio compared to age_lt5
age_rr <- age_rr %>%
  mutate(
        SE_var = (cum_upperCI - cum_lowerCI) / (2 * 1.96),
        SE_ref  = (ref_cumfit_disupper - ref_cumfit_dislower) / (2 * 1.96),
        SE_diff = sqrt(SE_var^2 + SE_ref^2),
            
        RR_ratio = exp(cumulative_fit - ref_cumfit_dis),
        RR_ratio_lower = exp((cumulative_fit - ref_cumfit_dislower) - 1.96 * SE_diff),
        RR_ratio_upper = exp((cumulative_fit - ref_cumfit_disupper) + 1.96 * SE_diff)
  )

# Plot
age_rr <- subset(age_rr, age_rr$lag=="lag3")
age_rr$GPSC <- gsub("_count","",age_rr$GPSC)
# categorize for facet
age_rr$type <- NA
age_rr$type[which(age_rr$GPSC%in%c("pcv7","pcv13","nvt"))] <- "Vaccine Status"
age_rr$type[which(age_rr$GPSC%in%c("age_lt6","age_6t14","age_15t64","age_18t64","age_gt65","female"))] <- "Demographic Group"
age_rr$type[which(age_rr$GPSC%in%c("Sero14","Sero19F","Sero23F","Sero4"))] <- "PCV7"
age_rr$type[which(age_rr$GPSC%in%c("Sero1","Sero19A","Sero3","Sero6A"))] <- "PCV13"
age_rr$type[which(age_rr$GPSC%in%c("Sero12F","Sero8"))] <- "NVT"
age_rr$type[which(age_rr$GPSC%in%c("other","mening","bact","WGS_PEN_SIR_Meningitis_R"))] <- "Disease Type"
age_rr$type <- factor(age_rr$type, levels = c("Demographic Group", "Disease Type", "Vaccine Status","PCV7","PCV13","NVT"))

age_rr$GPSC <- factor(age_rr$GPSC, levels = c("age_lt6","age_6t14","age_15t64","age_18t64","age_gt65","female","other","mening","bact","WGS_PEN_SIR_Meningitis_R",
"pcv7","pcv13","nvt","Sero14","Sero19F","Sero23F","Sero4","Sero1","Sero19A","Sero3","Sero6A","Sero12F","Sero8"),
                      labels = c(c("age <6","age 6-14","age 15-64","age 18-64","age 65+","female","other invasive disease","meningitis","bacteremia","penicillin resistant",
"PCV7","PCV13","NVT","Sero14","Sero19F","Sero23F","Sero4","Sero1","Sero19A","Sero3","Sero6A","Sero12F","Sero8") ))



prr <- ggplot(age_rr,aes(x = GPSC, y = RR_ratio, group = interaction(cov, type))) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "gray") +
  geom_point(size = 3) +
  scale_y_continuous(trans = "log10")+
  geom_errorbar(aes(ymin = RR_ratio_lower, ymax = RR_ratio_upper, group = interaction(cov,type)), width = 0.2) +
  labs( title = "",x = "Outcome", y = "Relative Risk Ratio\n(outcome/all disease)") +
  facet_wrap(.~type, nrow =1, scales = "free_x", strip.position = "bottom")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=45,hjust=1, size = 13), strip.text.x.bottom = element_text(size=13), axis.text = element_text(size=13),
        axis.title = element_text(size =13), strip.background = element_rect(fill = "white", color = NA))
age_rr[,c("GPSC","RR","RR_lower","RR_upper","RR_ratio","RR_ratio_lower","RR_ratio_upper")]
prr
  
pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/rrratio_suboutcomes.pdf", width = 12,height=4)
print(prr)
dev.off()

#### numbers for text
age_rr[which(age_rr$GPSC=="female" ),c("GPSC","RR","RR_lower","RR_upper","RR_ratio")]
age_rr[which(age_rr$GPSC=="age <6" ),c("GPSC","RR","RR_lower","RR_upper","RR_ratio")]
age_rr[which(age_rr$GPSC=="other invasive disease" ),c("GPSC","RR","RR_lower","RR_upper","RR_ratio")]
age_rr[which(age_rr$GPSC=="Sero14" ),c("GPSC","RR","RR_lower","RR_upper","RR_ratio", "RR_ratio_lower","RR_ratio_upper")]


data2<- fread(file="/home/sbelman/Documents/BRD/SouthAfrica/disease/SA_disease_point_base.csv",quote=FALSE, header = TRUE)
table(data2[which(data2$serotype=="23F")]$GPSC)/sum(table(data2[which(data2$serotype=="23F")]$GPSC))
table(data2[which(data2$serotype=="8")]$GPSC)/sum(table(data2[which(data2$serotype=="8")]$GPSC))
table(data2[which(data2$serotype=="8")]$GPSC)
table(data2[which(data2$serotype=="4")]$GPSC)/sum(table(data2[which(data2$serotype=="4")]$GPSC))
table(data2[which(data2$serotype=="4")]$GPSC)

age_rr[which(age_rr$RR_ratio_lower>0.99),c("GPSC","RR_ratio","RR_ratio_lower","RR_ratio_upper")]

```
##### text numbers for demographic distributions
```{r}
fits_age$cum_fitRR <- exp(fits_age$cumulative_fit)
fits_age$cumfit_lowerRR <- exp(fits_age$cum_lowerCI)
fits_age$cumfit_upperRR <- exp(fits_age$cum_upperCI)

fits_age[which(fits_age$GPSC=="age_gt65" & fits_age$var==50 & fits_age$lag_num == 3),c("GPSC","var","cum_fitRR","cumfit_lowerRR","cumfit_upperRR")]
fits_age[which(fits_age$GPSC=="age_lt6" & fits_age$var==100 & fits_age$lag_num == 3),c("GPSC","var","cum_fitRR","cumfit_lowerRR","cumfit_upperRR")]


fits_age[which(fits_age$cumfit_lowerRR>1&fits_age$predvar==50&fits_age$lag_num==3),c("GPSC","var","cum_fitRR","cumfit_lowerRR","cumfit_upperRR")]
age_rr[,c("GPSC","RR_ratio","RR_ratio_lower","RR_ratio_upper")]
```

#### Figure 4 and Figure S15: disease types distributions
```{r}
  fits_distypes_cum <- fits_distypes[which(fits_distypes$lag_num==3),]
  fits_distypes_cum$GPSC <- factor(fits_distypes_cum$GPSC, levels = c("bact_count","mening_count","other_count", "WGS_PEN_SIR_Meningitis_R_count"), labels = c("bact_count","mening_count","other_count", "penR_count"))
    # fits_distypes_cum$GPSC <- factor(fits_distypes_cum$GPSC, levels = c("bact_count","mening_count"))

  fits_distypes_cum$cum_fitRR <- exp(fits_distypes_cum$cumulative_fit)
    fits_distypes_cum$cumfit_lowerRR <- exp(fits_distypes_cum$cum_lowerCI)
  fits_distypes_cum$cumfit_upperRR <- exp(fits_distypes_cum$cum_upperCI)

  plot_dis <- ggplot(fits_distypes_cum)+
      geom_line(aes(x=predvar,y=exp(cumulative_fit_dis),group=interaction(lag_week,GPSC)), color="black", alpha=0.7, linetype = "dashed")+
      geom_line(aes(x=predvar,y=exp(cumulative_fit),group=interaction(lag_week,GPSC,cov), color=GPSC),alpha=0.7)+
      geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
      geom_ribbon(aes(x=predvar, ymin=exp(cum_lowerCI),ymax=exp(cum_upperCI), group=interaction(lag_week,GPSC,cov), fill = GPSC),alpha=0.1)+
      theme_bw()+
      xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
      ylab("Relative Risk")+
      # scale_y_continuous(trans="log10", limits = c(0.9,1.25), breaks = c(0.9, 1, 1.25))+
      scale_y_continuous(trans="log10")+
      scale_color_manual(labels = c("bact_count"="bacteremia","mening_count"="meningitis","other_count"="other invasive disease"
                                    # ),
                                    ,"penR_count" = "penicillin resistant"),
                         values = c("bact_count"="#D7191C","mening_count"="#2C7BB6","other_count"="#1A9641"
                                    # ))+
                                    ,"penR_count" = "purple"))+
      scale_fill_manual(labels = c("bact_count"="bacteremia","mening_count"="meningitis","other_count"="other invasive disease"
                                   # ),
                                   ,"penR_count" = "penicillin resistant"),
                        values =  c("bact_count"="#D7191C","mening_count"="#2C7BB6","other_count"="#1A9641"
                                    # ))+
                                    ,"penR_count" = "purple"))+
      ggtitle("PM2.5")+
      # facet_wrap(GPSC~. , ncol=3)+
      labs(color="Disease Type",fill="Disease Type")+
      theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
            strip.text = element_text(size=9), axis.text.x = element_text(angle = 45,hjust=1, size=13), 
            legend.position = "right")
    plot_dis
    ggsave(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/disease_type_pm2p5_cumulative.png", width = 5,height=3)
#     
pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/disease_type_pm2p5_cumulative.pdf", width = 5,height=3)
print(plot_dis)
dev.off()



   ### across lags
   fits_distypeslag <- fits_distypes[which(fits_distypes$lag_num<9),]
   fits_distypeslag$upperCI[which(fits_distypeslag$upperCI>0.1133287)] <- 0.1133287
     fits_distypeslag$GPSC <- factor(fits_distypeslag$GPSC, levels = c("bact_count","mening_count","other_count", "WGS_PEN_SIR_Meningitis_R_count"), labels = c("bact_count","mening_count","other_count", "penR_count"))
   plot_dislag <- ggplot(fits_distypeslag)+
      geom_line(aes(x=predvar,y=exp(fit),group=interaction(lag_week,GPSC,cov), color=GPSC),alpha=0.7)+
      geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
      geom_ribbon(aes(x=predvar, ymin=exp(lowerCI),ymax=exp(upperCI), group=interaction(lag_week,GPSC,cov), fill = GPSC),alpha=0.1)+
      # geom_line(aes(x=predvar,y=exp(fit_dis),group=interaction(lag_week,GPSC)), color="black", alpha=0.7)+
      # geom_ribbon(aes(x=predvar, ymin=exp(lowerCI_dis),ymax=exp(upperCI_dis), group=interaction(lag_week,GPSC)), fill = "black" ,alpha=0.1)+
      theme_bw()+
      xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
      ylab("Relative Risk")+
      # scale_y_continuous(trans="log10", limits = c(0.9,1.25), breaks = c(0.9, 1, 1.25))+
      scale_y_continuous(trans="log10", limits = c(0.75, 1.13), breaks = c(0.8,1,1.10))+
          scale_color_manual(labels = c("bact_count"="bacteremia","mening_count"="meningitis","other_count"="other invasive disease","penR_count" = "penicillin resistant"), 
                         values = c("bact_count"="#D7191C","mening_count"="#2C7BB6","other_count"="#1A9641"
                                    ,"penR_count" = "purple"))+
      scale_fill_manual(labels = c("bact_count"="bacteremia","mening_count"="meningitis","other_count"="other invasive disease","penR_count" = "penicillin resistant"),
                        values =  c("bact_count"="#D7191C","mening_count"="#2C7BB6","other_count"="#1A9641"
                                    ,"penR_count" = "purple"))+
      ggtitle("PM2.5")+
      facet_grid(GPSC~lag_num ,scales = "free_y")+
      labs(color="Disease Type",fill="Disease Type")+
      theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
            strip.text = element_text(size=9, angle=0), strip.text.y = element_text(size=9, angle=0), axis.text.x = element_text(angle = 45,hjust=1, size=13), 
            legend.position = "right")
   pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/disease_type_pm2p5_weekly.pdf", width = 13,height=4)
  print(plot_dislag)
  dev.off()
   
fits_distypes_cum[which(fits_distypes_cum$cumfit_lowerRR>0.9&fits_distypes_cum$predvar==50&fits_distypes_cum$lag_num==3),c("GPSC","var","cum_fitRR","cumfit_lowerRR","cumfit_upperRR")]

```


#### Figure S17: serotype distributions
```{r}
# Create a nested label based on pcv_type and serotype
nested_labels <- c(
  "Sero14"  = "PCV7:   Sero14",
  "Sero19F" = "PCV7:   Sero19F",
  "Sero23F" = "PCV7:   Sero23F",
  "Sero4"   = "PCV7:   Sero4",

  "Sero1"   = "PCV13:  Sero1",
  "Sero19A" = "PCV13:  Sero19A",
  "Sero3"   = "PCV13:  Sero3",
  "Sero6A"  = "PCV13:  Sero6A",

  "Sero8"   = "NVT:    Sero8",
  "Sero12F" = "NVT:    Sero12F"
)

fits_sero_cum <- fits_sero[which(fits_sero$lag_num==3),]
# Apply these labels to a new variable in your dataframe
fits_sero_cum$GPSC_nested <- nested_labels[as.character(fits_sero_cum$GPSC)]
fits_sero_cum$pcv_type <- factor(fits_sero_cum$pcv_type, levels = c("NVT","PCV13","PCV7"))

  fits_sero_cum$cum_fitRR <- exp(fits_sero_cum$cumulative_fit)
  fits_sero_cum$cumfit_lowerRR <- exp(fits_sero_cum$cum_lowerCI)
  fits_sero_cum$cumfit_upperRR <- exp(fits_sero_cum$cum_upperCI)

  
# Plot using the new nested label
plot_sero <- ggplot(fits_sero_cum) +
              geom_line(aes(x=predvar,y=exp(cumulative_fit_dis),group=interaction(lag_week,GPSC)), 
                        color="black", alpha=0.7, linetype = "dashed")+
              geom_line(aes(x = predvar, y = exp(cumulative_fit),
                            group = interaction(lag_week, GPSC, cov, pcv_type),
                            color = GPSC_nested), alpha = 0.7) +
              geom_ribbon(aes(x = predvar, ymin = exp(cum_lowerCI), ymax = exp(cum_upperCI),
                              group = interaction(lag_week, GPSC, cov, pcv_type),
                              fill = GPSC_nested), alpha = 0.1) +
              geom_hline(yintercept = 1, linetype = "dashed", color = "red", alpha = 0.6) +
              facet_wrap(. ~ pcv_type, ncol = 1) +
              labs(color = "Serotype", fill = "Serotype") +
              theme_bw() +
              theme(axis.text = element_text(size = 13),
                    axis.title = element_text(size = 13),
                    strip.text = element_text(size = 9),
                    axis.text.x = element_text(angle = 45, hjust = 1, size = 13),
                    legend.title = element_text(size = 12)) +
              xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')'))) +
              ylab("Relative Risk") +
              ggtitle("PM2.5") +
              scale_y_continuous(trans = "log10")

pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/serotype_pm2p5_cumulative.pdf", width = 4,height=10)
print(plot_sero)
dev.off()
plot_sero
ggsave(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/serotype_pm2p5_cumulative.png", width = 4,height=10)

############# lagged effects
fits_sero_week <- fits_sero[which(fits_sero$lag_num%in%c(0:8)),]
# Apply these labels to a new variable in your dataframe
fits_sero_week$GPSC_nested <- nested_labels[as.character(fits_sero_week$GPSC)]
fits_sero_week$GPSC <- factor(fits_sero_week$GPSC, levels = c("Sero12F","Sero8",
                                                              "Sero1","Sero19A","Sero3","Sero6A",
                                                              "Sero14","Sero19F","Sero23F","Sero4"))
fits_sero_week$upperCI[which(fits_sero_week$upperCI>=0.4054651)] <- 0.4054651
plot_sero_week <- ggplot(fits_sero_week) +
              # geom_line(aes(x=predvar,y=exp(cumulative_fit_dis),group=interaction(lag_week,GPSC)), 
                        # color="black", alpha=0.7, linetype = "dashed")+
              geom_line(aes(x = predvar, y = exp(fit),
                            group = interaction(lag_week, GPSC, cov, pcv_type),
                            color = GPSC_nested)) +
              geom_ribbon(aes(x = predvar, ymin = exp(lowerCI), ymax = exp(upperCI),
                              group = interaction(lag_week, GPSC, cov, pcv_type),
                              fill = GPSC_nested), alpha = 0.1) +
              geom_hline(yintercept = 1, linetype = "dashed", color = "red", alpha = 0.6) +
              facet_wrap(pcv_type ~ ., ncol = 3) +
              labs(color = "Serotype", fill = "Serotype") +
              theme_bw() +
              theme(axis.text = element_text(size = 13),
                    axis.title = element_text(size = 13),
                    strip.text = element_text(size = 9),
                    axis.text.x = element_text(angle = 45, hjust = 1, size = 13),
                    legend.title = element_text(size = 12)) +
              xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')'))) +
              ylab("Relative Risk") +
              facet_grid(GPSC~lag_num) + 
              # ggtitle("PM2.5") +
              scale_y_continuous(trans = "log10", limits = c(0.7, 1.5), breaks = c(0.8,1,1.12,1.5))
pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/serotype_pm2p5_weekly.pdf", width = 15,height=12)
print(plot_sero_week)
dev.off()

print(plot_sero_week)
ggsave(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/serotype_pm2p5_weekly.png", width = 15,height=12)
dev.off()

fits_sero_cum[which(fits_sero_cum$cumfit_lowerRR>0.999&fits_sero_cum$predvar==50&fits_sero_cum$lag_num==3),c("GPSC","var","cum_fitRR","cumfit_lowerRR","cumfit_upperRR")]
age_rr[which(age_rr$RR_ratio_lower>0.99),c("GPSC","RR_ratio","RR_ratio_lower","RR_ratio_upper")]


##### SEROTYPE 19f ALONE
plot19f <- ggplot(fits_sero_week[which(fits_sero_week$GPSC=="Sero19F"),]) +
                        # color="black", alpha=0.7, linetype = "dashed")+
              geom_line(aes(x = predvar, y = exp(fit),
                            group = interaction(lag_week, GPSC, cov, pcv_type)), color = "darkgreen") +
              geom_ribbon(aes(x = predvar, ymin = exp(lowerCI), ymax = exp(upperCI),
                              group = interaction(lag_week, GPSC, cov, pcv_type)),fill = "darkgreen", alpha = 0.1) +
              geom_hline(yintercept = 1, linetype = "dashed", color = "red", alpha = 0.6) +
              theme_bw() +
              theme(axis.text = element_text(size = 11),
                    axis.title = element_text(size = 12),
                    strip.text = element_text(size = 9),
                    axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
                    legend.title = element_text(size = 12)) +
              xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')'))) +
              ylab("Relative Risk") +
              facet_grid(.~lag_num) + 
              scale_y_continuous(trans = "log10", limits = c(0.7, 1.5), breaks = c(0.8,1,1.12,1.5))


### cumulative 19f
plot_seropcv7 <- ggplot(fits_sero_cum[which(fits_sero_cum$pcv_type=="PCV7"& fits_sero_cum$GPSC=="Sero19F"),]) +
              geom_line(aes(x=predvar,y=exp(cumulative_fit_dis),group=interaction(lag_week,GPSC)), 
                        color="black", alpha=0.7, linetype = "dashed")+
              geom_line(aes(x = predvar, y = exp(cumulative_fit),
                            group = interaction(lag_week, GPSC, cov, pcv_type)), color = "darkgreen", alpha = 0.7) +
              geom_ribbon(aes(x = predvar, ymin = exp(cum_lowerCI), ymax = exp(cum_upperCI),
                              group = interaction(lag_week, GPSC, cov, pcv_type)), fill="darkgreen", alpha = 0.1) +
              geom_hline(yintercept = 1, linetype = "dashed", color = "red", alpha = 0.6) +
              facet_grid(GPSC ~ pcv_type) +
              labs(color = "Serotype", fill = "Serotype") +
              theme_bw() +
              theme(axis.text = element_text(size = 13),
                    axis.title = element_text(size = 13),
                    strip.text = element_text(size = 9),
                    axis.text.x = element_text(angle = 45, hjust = 1, size = 13),
                    legend.title = element_text(size = 12)) +
              xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')'))) +
              ylab("Relative Risk") +
              ggtitle("PM2.5") +
              scale_y_continuous(trans = "log10")
```
#### Figure 4D and Figure S16 vt vs nvt distributions
```{r}
pcv_cols <- c("nvt"="#E75480","pcv7" ="#522081","pcv13"="darkgreen")

### plot cumulative
  fits_pcv_type_cum <- fits_pcv_type[which(fits_pcv_type$lag_num==3),]
  fits_pcv_type_cum$GPSC <- factor(fits_pcv_type_cum$GPSC, levels = c("nvt","pcv7","pcv13"))
  plot_pcv_cum <- ggplot(fits_pcv_type_cum)+
       geom_line(aes(x=predvar,y=exp(cumulative_fit_dis),group=interaction(lag_week,GPSC)), color="black", alpha=0.7, linetype = "dashed")+
      geom_line(aes(x=predvar,y=exp(cumulative_fit),group=interaction(lag_week,GPSC,cov), color=GPSC),alpha=0.7)+
      geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
      geom_ribbon(aes(x=predvar, ymin=exp(cum_lowerCI),ymax=exp(cum_upperCI), group=interaction(lag_week,GPSC,cov), fill = GPSC),alpha=0.1)+
      # geom_line(aes(x=predvar,y=exp(fit_dis),group=interaction(lag_week,GPSC)), color="black", alpha=0.7)+
      # geom_ribbon(aes(x=predvar, ymin=exp(lowerCI_dis),ymax=exp(upperCI_dis), group=interaction(lag_week,GPSC)), fill = "black" ,alpha=0.1)+
      scale_color_manual(values =pcv_cols )+
      scale_fill_manual(values =pcv_cols )+
      theme_bw()+
      xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
      ylab("Relative Risk")+
      # scale_y_continuous(trans="log10", limits = c(0.9,1.25), breaks = c(0.9, 1, 1.25))+
      scale_y_continuous(trans="log10")+
      ggtitle("PM2.5")+
      # facet_wrap(GPSC~. , ncol=3)+
      labs(color="Vaccine Type",fill="Vaccine Type")+
      theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
            strip.text = element_text(size=9), axis.text.x = element_text(angle = 45,hjust=1, size=13), 
            legend.position = "right")
    plot_pcv_cum
pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/pcv_pm2p5_cumulative.pdf", width = 5,height=3)
print(plot_pcv_cum)
dev.off()
plot_pcv_cum
ggsave(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/pcv_pm2p5_cumulative.png", width = 5,height=3)

    #### dlnm norm fit
    fit_disease <- dlnm_norm_fit[which(dlnm_norm_fit$cov == "pm2p5"),c("predvar","lag_num","fit_dis")]
    fit_disease$rr_fit_disnorm <- exp(fit_disease$fit_dis)
    fits_pcv_type2 <- left_join(fits_pcv_type,fit_disease[,c("predvar","lag_num","rr_fit_disnorm")],by=c("lag_num","predvar"))
    
    #### plot first 4 weeks
     fits_pcv_type_4 <- fits_pcv_type2[which(fits_pcv_type2$lag_num%in%0:8),]
  fits_pcv_type_4$GPSC <- factor(fits_pcv_type_4$GPSC, levels = c("nvt","pcv7","pcv13"), labels =  c("nvt","pcv7","pcv13"))
  plot_pcv_4 <- ggplot(fits_pcv_type_4)+
      geom_line(aes(x=predvar,y=rr_fit_disnorm,group=interaction(lag_week,GPSC)), color="black", alpha=0.7, linetype = "dashed")+
      geom_line(aes(x=predvar,y=exp(fit),group=interaction(lag_week,GPSC,cov), color=GPSC),alpha=0.7)+
      geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
      geom_ribbon(aes(x=predvar, ymin=exp(lowerCI),ymax=exp(upperCI), group=interaction(lag_week,GPSC,cov), fill = GPSC),alpha=0.1)+
      # geom_ribbon(aes(x=predvar, ymin=exp(lowerCI_dis),ymax=exp(upperCI_dis), group=interaction(lag_week,GPSC)), fill = "black" ,alpha=0.1)+
      scale_color_manual(values =pcv_cols )+
      scale_fill_manual(values =pcv_cols )+
      theme_bw()+
      xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
      ylab("Relative Risk")+
      # scale_y_continuous(trans="log10", limits = c(0.9,1.25), breaks = c(0.9, 1, 1.25))+
      scale_y_continuous(trans="log10")+
      ggtitle("PM2.5")+
      facet_wrap(lag_num~. , ncol=9)+
      labs(color="Vaccine Type",fill="Vaccine Type")+
      theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
            strip.text = element_text(size=9), axis.text.x = element_text(angle = 45,hjust=1, size=13), 
            legend.position = "right")
  plot_pcv_4
pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/pcv_pm2p5_8weeks.pdf", width = 12,height=3)
print(plot_pcv_4)
dev.off()
plot_pcv_cum
ggsave(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure4/pcv_pm2p5_cumulative.png", width = 12,height=3)

```

## Figure S6: CUMULATIVE PLOT for less important variables (SPEI, O3, PRLR, SFCWIND)
```{r}
space_time_str <- "weekly_adm2"
data_unscaled <- fread("/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm2_weekly_lag.csv")
data_unscaled_aggregate <- data_unscaled %>%
  group_by(district) %>%
  summarise(spei3 = mean(spei3_lag0,na.rm=T),
            spei6 = mean(spei6_lag0,na.rm=T),
            o3 = mean(o3_lag0,na.rm=T),
            sfcWind = mean(sfcWind_lag0,na.rm=T))
# SPEI
tmp <- subset(fit_list, fit_list$cov%in%c("spei3","spei6") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
# tmp$cov2 <- factor(tmp$cov2, levels = c("pm2p5","pm10","so2","tasmax","none"))
plot_spei_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov), color=cov),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov), fill=cov),alpha=0.3)+
    geom_rug(data= data_unscaled, aes(x = spei3_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  scale_color_manual(values = c("spei3" = "purple","spei6" ="darkred"))+
  scale_fill_manual(values = c("spei3" = "purple", "spei6"="darkred"))+
  # scale_linetype_manual(values = linetype_labels)+
  labs(color = "SPEI", fill = "SPEI")+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10", limits = c(0.4,1.1), breaks = c(0.4, 1, 1.1))+
    scale_y_continuous(trans="log10", limits = c(0.7,1.2), breaks = c(0.8, 1, 1.2))+
  ggtitle("SPEI")+
  xlab("SPEI")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "right")

##### PRECIPITATION
tmp <- subset(fit_list, fit_list$cov%in%c("prlrsum") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
# tmp$cov2 <- factor(tmp$cov2, levels = c("pm2p5","pm10","so2","tasmax","none"))
plot_prlrsum_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov), color=cov),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov), fill=cov),alpha=0.3)+
    geom_rug(data= data_unscaled, aes(x = prlrsum_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  scale_color_manual(values = c("prlrsum" = "#000080"))+
  scale_fill_manual(values = c("prlrsum" = "#000080"))+
  # scale_linetype_manual(values = linetype_labels)+
  labs(color = "", fill = "")+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10")+
  ggtitle("Precipitation (sum)")+
  xlab("millimeters")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "none")

##### PRECIPITATION
tmp <- subset(fit_list, fit_list$cov%in%c("prlrmean") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
# tmp$cov2 <- factor(tmp$cov2, levels = c("pm2p5","pm10","so2","tasmax","none"))
plot_prlrmean_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov), color=cov),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov), fill=cov),alpha=0.3)+
    geom_rug(data= data_unscaled, aes(x = prlrmean_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  scale_color_manual(values = c("prlrmean" = "#6699CC","prlrsum" = "#000080"))+
  scale_fill_manual(values = c("prlrmean" = "#6699CC", "prlrsum" = "#000080"))+
  # scale_linetype_manual(values = linetype_labels)+
  labs(color = "", fill = "")+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10")+
  ggtitle("Precipitation (mean)")+
  xlab("millimeters")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "none")


##### WIND
tmp <- subset(fit_list, fit_list$cov%in%c("sfcWind") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
# tmp$cov2 <- factor(tmp$cov2, levels = c("pm2p5","pm10","so2","tasmax","none"))
plot_wind_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov)),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov)),alpha=0.3)+
    geom_rug(data= data_unscaled, aes(x = sfcWind_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  labs(color = "", fill = "")+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10")+
  ggtitle("Wind")+
  xlab("km/hr")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "right")


##### OZONE
tmp <- subset(fit_list, fit_list$cov%in%c("o3") & fit_list$lag_num %in% c(3) & fit_list$space_time==space_time_str)
# tmp$cov2 <- factor(tmp$cov2, levels = c("pm2p5","pm10","so2","tasmax","none"))
plot_o3_cum <- ggplot(tmp)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_week,space_time, cov)),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_week,space_time,cov)),alpha=0.3)+
    geom_rug(data= data_unscaled, aes(x = o3_lag0), sides = "b", alpha = 0.03, length = unit(0.05, "npc"), color = "grey40") +
  # scale_color_manual(values = c("prlrmean" = "blue"))+
  # scale_fill_manual(values = c("prlrmean" = "blue"))+
  # scale_linetype_manual(values = linetype_labels)+
  labs(color = "", fill = "")+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10", breaks = c(0.8,1, 1.3))+
  ggtitle("Ozone")+
  xlab("Concentration")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "right")

### save each individually
pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/prlrmean_cumulative_additive_sa_",space_time_str,".pdf"),width  =4.5,height=3.5)
print(plot_prlrmean_cum)
dev.off()
pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/prlrsum_cumulative_additive_sa_",space_time_str,".pdf"),width  =4.5,height=3.5)
print(plot_prlrsum_cum)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/spei_cumulative_additive_sa_",space_time_str,".pdf"),width  =4.5,height=3.5)
print(plot_spei_cum)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/wind_cumulative_additive_sa_",space_time_str,".pdf"),width  =4.5,height=3.5)
print(plot_wind_cum)
dev.off()

pdf(paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/o3_cumulative_additive_sa_",space_time_str,".pdf"),width  =4.5,height=3.5)
print(plot_o3_cum)
dev.off()
```

### TEXT SECTION
### Print the numbers I need from Figure 3
```{r}
### particulate matter numbers PM2.5
tmppm2[which(tmppm2$var==50&tmppm2$lag=="lag3"),c("cov2","var","cumrr","cumlowerCI_rr","cumupperCI_rr")]

cumspms2 <- tmppm2[which( tmppm2$lag=="lag3" & tmppm2$cov == "pm2p5" & tmppm2$cov2 == "absh_lag3 + tas_lag0"),c("cov2","var","cumrr","cumlowerCI_rr","cumupperCI_rr")]
cumspms2$perc_increase_above40 <- (cumspms2$cumrr / cumspms2$cumrr[(cumspms2$var == 40 & cumspms2$cov2 == "absh_lag3 + tas_lag0")] - 1) * 100
cumspms2[which(cumspms2$cov2=="none"),]
cumspms2[which(cumspms2$cov2=="cumspms2"),]
## effect size at key concentrations
cumspms2[which(cumspms2$var=="50"),]
cumspms2[which(cumspms2$var=="100"),]
## maximum effect size
tmppm250 <- tmppm2[which(tmppm2$var==50 & tmppm2$cov2=="absh_lag3 + tas_lag0")]

testplot <- ggplot(tmppm2[which(tmppm2$cov2=="absh_lag3 + tas_lag0"),])+
  geom_line(aes(x=var,y=rr,group=interaction(lag_week,space_time, cov2)),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(lag_week,space_time,cov2)),alpha=0.5)+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.85,1.15), breaks = c(0.85, 1, 1.15))+
  ggtitle("PM 2.5")+
  facet_wrap(.~lag_week, nrow = 1)+
  xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13))

tmppm25 <- tmppm2[tmppm2$cov2 == "absh_lag3 + tas_lag0" & tmppm2$lag_num<5, ]
max_rr_row <- tmppm25[which.max(tmppm25$lowerCI), ]
cat("Lag week with maximum relative risk:", max_rr_row$lag_week, "\n")
cat("Concentration at peak:", max_rr_row$var, "\n")
cat("Maximum relative risk:", max_rr_row$fit, "\n")
tmppm2[which(tmppm2$cov == "pm2p5"& tmppm2$cov2=="absh_lag3 + tas_lag0" & tmppm2$var==100),c("cov2","lag","var","cumrr","rr","lowerCI_rr","upperCI_rr")]


### particulate matter numbers PM10
cumspms10 <- tmppm1[which( tmppm1$lag=="lag3" & tmppm1$cov == "pm10") &tmppm1$cov2=="absh_lag3 + tas_lag0",c("cov","cov2","var","cumrr","cumlowerCI_rr","cumupperCI_rr")]
cumspms10$perc_increase_above75 <- (cumspms10$cumrr / cumspms10$cumrr[(cumspms10$var == 75 & cumspms10$cov2 == "absh_lag3 + tas_lag0")] - 1) * 100
cumspms10[which(cumspms10$cov2=="none"),]
cumspms10[which(cumspms10$var=="100"),]
tmppm1[which(tmppm1$cov == "pm10"& tmppm1$cov2=="absh_lag3 + tas_lag0" & tmppm1$var==150),c("cov2","lag","var","cumrr","rr","lowerCI_rr","upperCI_rr")]

### particulate matter numbers SO2
cumsso2 <- tmpso2[which( tmpso2$lag=="lag3" & tmpso2$cov == "so2"& tmpso2$cov2=="absh_lag3 + tas_lag0"),c("cov2","var","cumrr","cumlowerCI_rr","cumupperCI_rr")]
cumsso2$perc_increase_above125 <- (cumsso2$cumrr / cumsso2$cumrr[(cumsso2$var == 130 & cumsso2$cov2 == "absh_lag3 + tas_lag0")] - 1) * 100
cumsso2[which(cumsso2$cov2=="none"),]
cumsso2[which(cumsso2$var=="100"),]
 tmpso2[which(tmpso2$cov == "so2"& tmppm1$cov2=="absh_lag3 + tas_lag0" & tmpso2$var==150),c("cov2","lag","var","rr","lowerCI_rr","upperCI_rr")]

### humidity numbers
## low humidity
tmphurs[which(tmphurs$rr==max(tmphurs$rr))]
highhursrisk <- tmphurs[which(tmphurs$cumlowerCI_rr>1 & tmphurs$cov2%in%c("none","pm2p5")),c("lag","cov2","var","rr","lowerCI_rr","upperCI_rr","cumrr","cumlowerCI_rr","cumupperCI_rr")]
highhursrisk[highhursrisk$var>35&highhursrisk$var<40]
tmphurs[which(tmphurs$var>39&tmphurs$var<40 & tmphurs$lag=="lag3"),c("lag","cov2","var","cumrr","cumlowerCI_rr","cumupperCI_rr")]
tmphurs[which(tmphurs$cumlowerCI_rr>1 & tmphurs$lag=="lag3"),c("var")]
### high humidity
tmphurs[which(tmphurs$var>74& tmphurs$var<76 & tmphurs$lag=="lag3"),c("cov2","var","lag","cumrr","cumlowerCI_rr","cumupperCI_rr")]
tmphurs[which(tmphurs$cumlowerCI_rr<1 &tmphurs$var>74 & tmphurs$lag=="lag3"),c("cov2","var","lag","cumrr","cumlowerCI_rr","cumupperCI_rr")]

tmphurscum <- tmphurs[which( tmphurs$lag=="lag3"),c("cov2","var","lag","cumrr","cumlowerCI_rr","cumupperCI_rr")]
tmphurscum[which(tmphurscum$cumupperCI_rr<1 & tmphurscum$var>40),]

### absolute humidity numbers
tmpabsh[which(max(tmpabsh$cumulative_fit)==tmpabsh$cumulative_fit & tmpabsh$lag=="lag3"),]
## low humidity
tmpabsh[which(tmpabsh$var>5.7&tmpabsh$var<5.8 & tmpabsh$lag=="lag3"),c("cov2","var","cumrr","cumlowerCI_rr","cumupperCI_rr")]

tmpabsh[which(tmpabsh$var>=4.0&tmpabsh$var<5.8 & tmpabsh$lag%in%c("lag0","lag1","lag2","lag3")),c("lag","cov2","var","rr","lowerCI_rr","upperCI_rr","cumrr","cumlowerCI_rr","cumupperCI_rr" )]
tmpabsh[which(tmpabsh$cumlowerCI_rr>1 & tmpabsh$lag=="lag3"),c("var")]


tmptas[which(tmptas$cov=="tasmin"&tmptas$lag=="lag3"&tmptas$cumlowerCI_rr>1),c("cov","var","cumrr","cumlowerCI_rr","cumupperCI_rr")]

tmptas[which(tmptas$cov=="tas" & tmptas$var==24),c("cov","var","lag","rr","lowerCI_rr","upperCI_rr")]

tmptas[which(tmptas$cov=="tasmax"&tmptas$var==35),c("cov","lag","var","rr","lowerCI_rr","upperCI_rr")]
```


## Scripts for visualizing Main Figure 5
#### Figure 5C: plot GPSC21 19F and pull numbers for manuscript
```{r}
time = "weekly"
space = "adm2"
gpsc_results <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/gpsc_results_fits_",time,"_",space,"_allGPSCs_propprov_bivariable_8_2019_templag0humlag3.csv"))
gpsc_results$cov2 <- "absh_lag3"
nointer <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/nointeraction_results_fits_",time,"_",space,"_bivariable_8_2019_templag0humlag3.csv"))
gpsc_results <- rbind(gpsc_results,nointer)
gpsc21 <- gpsc_results[which(gpsc_results$GPSC%in%c("GPSC21") & gpsc_results$interaction_level%in%c("high","none") & gpsc_results$covariate=="pm2p5_lag0")]
nointer <- gpsc_results[which(gpsc_results$GPSC%in%c("no_interaction") & gpsc_results$interaction_level%in%c("high","none") & gpsc_results$covariate=="pm2p5_lag0")]
ggplot(gpsc21) + 
  geom_line(aes(x=var, y = fit, group = interaction(lag_num)))+
  geom_ribbon(aes(x = var, ymin =lowerCI, ymax=upperCI, group = lag_num), alpha = 0.3)+ 
  geom_line(data= nointer, aes(x=var, y = fit, group = lag_num), color = "blue", linetype ="dotted") + 
  facet_grid(. ~lag_num, scales = "free")

gpsc21[which(gpsc21$lag_num==0 & gpsc21$var==50),]
exp(gpsc21[which(gpsc21$lag_num==1 & gpsc21$var==50),c("fit","lowerCI","upperCI")])
exp(nointer[which(nointer$lag_num==1 & nointer$var==50),c("fit","lowerCI","upperCI")])
```

#### Figure 5A: Proportion of GPSC 21 over space and time
```{r}
data2<- fread(file="/home/sbelman/Documents/env_sa_manuscript/input_datasets/disease/SA_disease_point_base.csv",quote=FALSE, header = TRUE)
# Convert date to week start (ISO week)
data2 <- data2 %>% 
  mutate(week = floor_date(date, "week"))
# Summarise total sequenced & GPSC21 count per province per week
      weekly_data <- data2 %>%
        filter(year > 2004 & year<2020) %>%
        group_by(week, province) %>%
        summarise(
          total_seq = sum(sequenced, na.rm = TRUE),
          gpsc21_seq = sum(ifelse(GPSC == "21", sequenced, 0), na.rm = TRUE),
          gpsc1_seq = sum(ifelse(GPSC == "1", sequenced, 0), na.rm = TRUE),
          .groups = "drop"
        ) %>%
        mutate(prop_gpsc21 = gpsc21_seq / total_seq,
               prop_gpsc1 = gpsc1_seq / total_seq)
      
      ### make sure 0's and NAs are different
      weekly_data2 <- weekly_data %>%
        mutate(prop_gpsc21plot = ifelse(total_seq == 0, NA, prop_gpsc21),
               prop_gpsc1plot = ifelse(total_seq == 0, NA, prop_gpsc1))
      
      plotgpsc21 <- ggplot(weekly_data2, aes(x = week, y = prop_gpsc21plot, size = prop_gpsc21plot, colour = prop_gpsc21plot)) +
          geom_point(alpha = 0.8) +
          scale_colour_gradientn(colours = c("#d6e9f9", "#4a7bd0", "#8b2f8c", "#d7191c"),values = scales::rescale(c(0, 0.25, 0.6, 1)),limits = c(0, 1),name = "Proportion GPSC21 per week")+
          scale_size(range = c(1, 8),limits = c(0, 1),name = "Proportion GPSC21") +
          scale_x_date(date_breaks = "5 years", date_labels = "%Y") +
          facet_wrap(~ province, ncol = 3, strip.position = "top") +
          labs(x = "Week", y = "Proportion GPSC21 (per week)") +
          theme_classic(base_size = 13) +
          ylim(0,1.3)+
          theme(
            strip.background = element_blank(),
            axis.text.x = element_text(angle = 45, hjust = 1, size =12),
            legend.box = "vertical",
            legend.position = "right"
          ) +
          guides(colour = guide_legend(override.aes = list(size = c(3, 6, 9, 12, 15))),size = "none")
      
      
pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure5/proportionsGPSC21.pdf", width = 8,height=4)
print(plotgpsc21)
dev.off()

print(plotgpsc21)
ggsave(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure5/proportionsGPSC21.png", width = 8,height=4)
dev.off()
```
### Figure 5B: proportion of GPSCs for each serotype 19f, 12f, and 6a
```{r}
data2<- fread(file="/home/sbelman/Documents/env_sa_manuscript/input_datasets/disease/SA_disease_point_base.csv",quote=FALSE, header = TRUE)
highlighted_gpscs <- c("1", "21", "26", "56", "13", "41")
gpsc_colors <- c(
  "1" = "pink",
  "21" = "#377EB8",
  "26" = "#4DAF4A",
  "56" = "darkblue",
  "13" = "#FF7F00",
  "41" = "purple",
  "Other" = "grey80"
)
#### proportion 19F
tmpdf <- data2[which(data2$serotype=="19F"),]
gpsc_year_counts <- tmpdf %>%
  group_by(year, GPSC) %>%
  summarise(n = n(), .groups = "drop")
gpsc_year_props <- gpsc_year_counts %>%
     mutate(GPSC_highlight = ifelse(GPSC %in% highlighted_gpscs, GPSC, "Other")) %>%
    filter(!is.na(GPSC), GPSC!="_", GPSC!="") %>%
  group_by(year) %>%
  mutate(total = sum(n),
         proportion = n / total)
p19f <- ggplot(gpsc_year_props, aes(x = factor(year), y = proportion, fill = GPSC_highlight)) +
  geom_bar(stat = "identity") +
  labs(title = "",
       x = "Year", y = "Proportion of 19F",
       fill = "GPSC") +
  theme_minimal() +
    theme(axis.text.x = element_text(angle=90,hjust=1))+
  scale_fill_manual(values = gpsc_colors)  # Or any palette you prefer

#### proportion 12F
tmpdf <- data2[which(data2$serotype=="12F"),]
gpsc_year_counts <- tmpdf %>%
  group_by(year, GPSC) %>%
  summarise(n = n(), .groups = "drop")
gpsc_year_props <- gpsc_year_counts %>%
     mutate(GPSC_highlight = ifelse(GPSC %in% highlighted_gpscs, GPSC, "Other")) %>%
    filter(!is.na(GPSC), GPSC!="_", GPSC!="") %>%
  group_by(year) %>%
  mutate(total = sum(n),
         proportion = n / total)
p12f <- ggplot(gpsc_year_props, aes(x = factor(year), y = proportion, fill = GPSC_highlight)) +
  geom_bar(stat = "identity") +
  labs(title = "",
       x = "Year", y = "Proportion of 12F",
       fill = "GPSC") +
  theme_minimal() +
    theme(axis.text.x = element_text(angle=90,hjust=1))+
  scale_fill_manual(values = gpsc_colors)  # Or any palette you prefer

#### proportion 6A

tmpdf <- data2[which(data2$serotype=="6A"),]
gpsc_year_counts <- tmpdf %>%
  group_by(year, GPSC) %>%
  summarise(n = n(), .groups = "drop")
gpsc_year_props <- gpsc_year_counts %>%
   mutate(GPSC_highlight = ifelse(GPSC %in% highlighted_gpscs, GPSC, "Other")) %>%
    filter(!is.na(GPSC), GPSC!="_", GPSC!="") %>%
  group_by(year) %>%
  mutate(total = sum(n),
         proportion = n / total)

p6A <- ggplot(gpsc_year_props, aes(x = factor(year), y = proportion, fill = GPSC_highlight)) +
  geom_bar(stat = "identity") +
  labs(title = "",
       x = "Year", y = "Proportion of 6A",
       fill = "GPSC") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=90,hjust=1))+
  scale_fill_manual(values = gpsc_colors)  # Or any palette you prefer

pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure5/proportionserotype_selectGPSCs_pm2p5.pdf", width = 4,height=6)
print(p19f/p12f/p6A)
dev.off()

pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/figure5/proportionserotype_GPSCswith19F_pm2p5.pdf", width = 4,height=3)
print(p19f)
dev.off()
```

## Figure S20B : Figure 4 associated
#### cumulative dose response curves for all GPSCs
```{r}
data2<- fread(file="/home/sbelman/Documents/env_sa_manuscript/input_datasets/disease/SA_disease_point_base.csv",quote=FALSE, header = TRUE)
dtgpsc <- data.table(table(data2$GPSC))[which(data.table(table(data2$GPSC))$N>100 | data.table(table(data2$GPSC))$V1%in%c("8","41"))]
gpsc_vec <- c(paste0("GPSC",dtgpsc[order(-N)]$V1))

space = "adm1"
tmps <- subset(allfits, allfits$cov== "pm2p5" & allfits$interaction_level %in% c("low","high","none") & allfits$lag_num==3 & allfits$space_time==paste0("weekly_",space))


no_int <- tmps %>% filter(GPSC == "no_interaction" & cov2=="absh_lag3") %>%
  rename_with(~paste0(., "_noint"), c("cumulative_fit", "cum_lowerCI", "cum_upperCI"))

main_gpsc <- tmps %>% filter(GPSC != "no_interaction")

# Step 2: Join them back by cov, var, and space_time
allfits_pivoted <- main_gpsc %>%
  left_join(no_int %>% 
              dplyr::select(space_time, cov, var,lag,cumulative_fit_noint, cum_lowerCI_noint, cum_upperCI_noint),
            by = c("var","cov","space_time","lag"))

## test
ggplot(allfits_pivoted[which(allfits_pivoted$interaction_level=="high"),])+
  geom_line(aes(x=var, y= cumulative_fit_noint, group = interaction(GPSC,cov)))
allfits_pivoted[which(allfits_pivoted$interaction_level=="high"&allfits_pivoted$GPSC=="GPSC1"&allfits_pivoted$predvar==50),]
##
## reorder
allfits_pivoted$GPSC <- factor(allfits_pivoted$GPSC, levels = gpsc_vec)
p14 <- ggplot(allfits_pivoted)+
  geom_line(aes(x=var,y=exp(cumulative_fit), group=interaction(cov, GPSC, interaction_level,lag),  color=interaction_level))+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_line(aes(x = var, y = exp(cumulative_fit_noint), group=interaction(cov, lag,interaction_level,GPSC)), color = "black", linetype="solid") +
  geom_ribbon(aes(x=var, ymin=exp(cum_lowerCI),ymax=exp(cum_upperCI), group=interaction(cov, GPSC, interaction_level,lag), fill=interaction_level),alpha=0.2)+
  geom_ribbon(aes(x=var, ymin=exp(cum_lowerCI_noint) ,ymax= exp(cum_upperCI_noint), group=interaction(cov,lag, GPSC)),fill="black",alpha=0.4)+
  # scale_color_manual(values = c("high"="red","low"="blue", "med" = "darkgreen")) +
  # scale_fill_manual(values = c("high"="red","low"="blue", "med" = "darkgreen")) +
   scale_color_manual(values = c("high"="red","low"="blue")) +
  scale_fill_manual(values = c("high"="red","low"="blue")) +
  # xlab(cov_names_labels[c])+
  scale_y_continuous(trans="log10")+
  theme_bw()+
  ylab("Relative Risk")+
  facet_wrap(GPSC~.,scales="free")+
  labs(fill="GPSC Prevalence", color = "GPSC Prevalence")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), strip.text = element_text(size=13), axis.text.x = element_text(angle=45, hjust = 1),
        strip.text.y = element_text(angle=0))
p14
pdf(file=paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure5/cumulative_effect_pm2p5",space,".pdf"), width = 8,height=7)
print(p14)
dev.off()
```
##### Figure S20A: 8 week lagged dose response curves
```{r}
space = "adm1"
tmps <- subset(allfits, allfits$cov== "pm2p5" & allfits$interaction_level %in% c("low","high","none") & allfits$space_time==paste0("weekly_",space))
# tmps <- subset(allfits, allfits$cov == "pm10" & allfits$interaction_level %in% c("low","none") & allfits$lag_num==3 & allfits$space_time=="weekly_adm2")
main_gpsc <- tmps %>% filter(GPSC != "no_interaction")
# 
no_int <- tmps %>% filter(GPSC == "no_interaction" & cov2 == "absh_lag3") %>%
  rename_with(~paste0(., "_noint"), c("fit", "lowerCI", "upperCI"))
# 
main_gpsc <- tmps %>% filter(GPSC != "no_interaction")
# 
# # Step 2: Join them back by cov, var, and space_time
allfits_pivoted <- main_gpsc %>%
  left_join(no_int %>%
              dplyr::select(space_time, cov, var,lag,fit_noint, lowerCI_noint, upperCI_noint),
            by = c("var","cov","space_time","lag"))

allfits_pivoted$lagweek <- paste0("week ",allfits_pivoted$lag_num)
allfits_pivoted$lagweek <- factor(allfits_pivoted$lagweek, levels = c("week 0", "week 1","week 2","week 3","week 4","week 5","week 6","week 7","week 8"))
allfits_pivoted$GPSC <- factor(allfits_pivoted$GPSC, levels = gpsc_vec)

p14 <- ggplot(allfits_pivoted)+
  geom_line(aes(x=var,y=exp(fit), group=interaction(cov, GPSC, interaction_level,lagweek),  color=interaction_level))+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=exp(lowerCI),ymax=exp(upperCI), group=interaction(cov, GPSC, interaction_level,lagweek), fill=interaction_level),alpha=0.2)+
    # geom_line(aes(x = var, y = exp(fit_noint), group=interaction(cov, GPSC, interaction_level,lagweek)), color = "black", linetype="solid") +
  scale_color_manual(values = c("high"="red","low"="blue", "med" = "darkgreen")) +
  scale_fill_manual(values = c("high"="red","low"="blue", "med" = "darkgreen")) +
  # ylim(0,3.5)+
  # xlab(cov_names_labels[c])
        scale_y_continuous(trans="log10")+
  theme_bw()+
  ylab("Relative Risk")+
  facet_grid(GPSC~lagweek,scales="free_x")+
  labs(fill="GPSC Prevalence", color = "GPSC Prevalence")+
  theme(axis.text = element_text(size=12),axis.title=element_text(size=13), strip.text = element_text(size=13), axis.text.x = element_text(angle = 45, hjust = 1), 
        strip.text.y = element_text(angle=0))

pdf(file= paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure5/lagged_effect_pm2p5_",space,".pdf"), width = 10,height=10)
print(p14)
dev.off()


#### select GPSCs
p15 <- ggplot(allfits_pivoted[which(allfits_pivoted$GPSC%in%c("GPSC14","GPSC21","GPSC26")),])+
  geom_line(aes(x=var,y=exp(fit), group=interaction(cov, GPSC, interaction_level,lagweek),  color=interaction_level))+
      # geom_line(aes(x = var, y = exp(fit_noint), group=interaction(cov, GPSC, interaction_level,lagweek)), color = "black", linetype="solid") +
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=exp(lowerCI),ymax=exp(upperCI), group=interaction(cov, GPSC, interaction_level,lagweek), fill=interaction_level),alpha=0.2)+
  scale_color_manual(values = c("high"="red","low"="blue", "med" = "darkgreen")) +
  scale_fill_manual(values = c("high"="red","low"="blue", "med" = "darkgreen")) +
  # ylim(0,3.5)+
  # xlab(cov_names_labels[c])+
          scale_y_continuous(trans="log10")+
  theme_bw()+
  ylab("Relative Risk")+
  facet_grid(GPSC~lagweek,scales="free_x")+
  labs(fill="GPSC Prevalence", color = "GPSC Prevalence")+
  theme(axis.text = element_text(size=10),axis.title=element_text(size=13), strip.text = element_text(size=13),
        strip.text.y = element_text(angle=0))
p15

pdf(file=paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure5/lagged_effect_selectGPSCs_pm2p5_",space,".pdf"), width = 10,height=3)
print(p15)
dev.off()

#### serotype 19F types
p16 <- ggplot(allfits_pivoted[which(allfits_pivoted$GPSC%in%c("GPSC1","GPSC21")),])+
  geom_line(aes(x=var,y=exp(fit), group=interaction(cov, GPSC, interaction_level,lagweek),  color=interaction_level))+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
        # geom_line(aes(x = var, y = exp(fit_noint), group=interaction(cov, GPSC, interaction_level,lagweek)), color = "black", linetype="dotted") +
  geom_ribbon(aes(x=var, ymin=exp(lowerCI),ymax=exp(upperCI), group=interaction(cov, GPSC, interaction_level,lagweek), fill=interaction_level),alpha=0.2)+
  scale_color_manual(values = c("high"="red","low"="blue", "med" = "darkgreen")) +
  scale_fill_manual(values = c("high"="red","low"="blue", "med" = "darkgreen")) +
  # ylim(0,3.5)+
  # xlab(cov_names_labels[c])+
          scale_y_continuous(trans="log10")+

  theme_bw()+
  ylab("Relative Risk")+
  xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+
  facet_grid(GPSC~lagweek,scales="free_x")+
  labs(fill="GPSC Prevalence", color = "GPSC Prevalence")+
  theme(axis.text = element_text(size=10),axis.title=element_text(size=13), strip.text = element_text(size=13),
        strip.text.y = element_text(angle=0), axis.text.x = element_text(angle=45, hjust =1))
p16

pdf(file= paste0("/home/sbelman/Documents/env_sa_manuscript/figures/figure5/lagged_effect_GPSCswith19F_pm2p5_",space,".pdf"), width = 10,height=3)
print(p16)
dev.off()
# 
# table(data2[which(data2$GPSC=="26"),]$serotype) ## serotype 12F (NVT) similar to on average (peak around 3-4 weeks)
# table(data2[which(data2$GPSC=="21"),]$serotype) ## serotype 19F (PCV7)  hardly any risk overall
# table(data2[which(data2$GPSC=="14"),]$serotype) ## serotype 23F (PCV7) acute risks for serotype 23F

```










# Scripts for visualizing supplemental figures
## FIGURE S1A-B: case counts weekly monthly and by province
```{r}
data<-fread("/home/sbelman/Documents/env_sa_manuscript/input_datasets/disease/SA_disease_point.csv")
weekly_data <- data %>%
  group_by(date,epi_year, epi_week, district) %>%
  summarise(weekly_value = sum(disease), .groups = 'drop')

data<-data.table(data)
dataggr <- data[, list(cases=sum(disease)), by=date]
dataaggr <- data[which(data$epi_year<2025),]
weekly_count <- ggplot(dataggr)+
  geom_line(aes(x=date,y=cases))+
  theme_classic()+
  xlab("Year")+
  ylab("Weekly Case Count")+
  scale_x_date(
    breaks = as.Date(c("2005-01-01", "2010-01-01", "2015-01-01", "2020-01-01")),
    labels = date_format("%Y")
  ) +
  theme(axis.text=element_text(size=17),axis.title = element_text(size=17),
        axis.text.x = element_text(angle=90))
ggsave("/home/sbelman/Documents/env_sa_manuscript/figures/figure1/country_daily.png",width=10,height=8)
####### Monthly Plot

#### monthly 
# Aggregate data to calculate monthly disease counts
monthly_data <- data %>%
  # Extract year and month from the date
  mutate(year = format(date, "%Y"),
         month = format(date, "%m")) %>%
  # Group by year, month, and district (optional)
  group_by(year, month) %>%
  # Summarize to get the monthly total of disease cases
  summarise(monthly_cases = sum(disease), .groups = 'drop')
# Combine year and month into a new column for plotting
monthly_data <- monthly_data %>%
  mutate(year_month = as.Date(paste(year, month, "15", sep = "-")))
# Convert to data.table (if needed)
monthly_data <- data.table(monthly_data)
# Plot the monthly case counts over time
month_count <- ggplot(monthly_data) +
  geom_line(aes(x = year_month, y = monthly_cases)) +
  theme_classic() +
  xlab("Year") +
  ylab("Monthly Case Count") +
  theme(axis.text = element_text(size = 17), 
        axis.title = element_text(size = 17),
        axis.text.x = element_text(angle = 90, hjust = 1),legend.position = "none")

#### province 
# Aggregate data to calculate monthly disease counts
color_labels = c("Eastern Cape" = "#008080", "Free State" = "#D75F00", "Gauteng" = "#967bb6" , "KwaZulu-Natal"= "#FFCCCB",
                     "Limpopo" = "#228B22", "Mpumalanga" = "#FFC000", "North West" = "#8B5A2B", "Northern Cape" = "#606060", "Western Cape" = "#1A2D6D" )
province_data <- data %>%
  # Extract year and month from the date
  mutate(year = format(date, "%Y"),
         month = format(date, "%m")) %>%
  # Group by year, month, and district (optional)
  group_by(province,year,month) %>%
  # Summarize to get the monthly total of disease cases
  summarise(province_cases = sum(disease), .groups = 'drop')
# Combine year and month into a new column for plotting
province_data <- province_data %>%
  mutate(year_month = as.Date(paste(year, month, "15", sep = "-")))
# Convert to data.table (if needed)
province_data <- data.table(province_data)
# Plot the monthly case counts over time
prov_count <- ggplot(province_data) +
  geom_line(aes(x = year_month, y = province_cases, color = province)) +
  scale_color_manual(values= color_labels )+ 
  theme_classic() +
  xlab("Year") +
  ylab("Monthly Case Count") +
  theme(axis.text = element_text(size = 17), 
        axis.title = element_text(size = 17),
        axis.text.x = element_text(angle = 90, hjust = 1),legend.position = "right", legend.title = element_blank())

data<-fread(file="/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm2_monthly_lag_sc.csv")
color_labels2 <- color_labels
names(color_labels2) <- gsub(" ","_",names(color_labels))
intplot <- ggplot(data) + 
  geom_line(aes(x=date,y=ART_coverage,group=NAME_1,color = NAME_1))+
  geom_line(aes(x=date,y=PCV_coverage_3dose, linetype="PCV Coverage"),color="black")+
  scale_color_manual(values=color_labels2)+
  theme_classic()+
  xlab("Year")+
  ylab("Coverage") +
  ylim(0,1)+
  labs(color="Antiretroviral Therapy\n(Province)",linetype="Pneumococcal Vaccine\n(National)") +
  theme(axis.text = element_text(size=17),axis.title = element_text(size=17),axis.text.x = element_text(angle=90))
pdf("/home/sbelman/Documents/env_sa_manuscript/figures/weeklymonthlyprov_casecount.pdf", width=12,height=10)
(weekly_count + month_count) /(prov_count+intplot) 
dev.off()
# (weekly_count + month_count) /(prov_count+intplot) + 
#   patchwork::plot_annotation(tag_levels = 'a')&
#   theme(plot.tag = element_text(size = 25))
# ggsave("/home/sbelman/Documents/env_sa_manuscript/figures/weeklymonthlyprov_casecount.png", width=12,height=4)

```
## FIGURE 3 TEXT SECTION ATTRIBUTABLE FRACTION
```{r}
data <-fread(file="/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm2_weekly_lag_sc.csv")
data_unscaled <- fread(file="/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm2_weekly_lag.csv")
## subset by Johnannesburg
dataG <- subset(data, data$adm2_name == "City of Cape Town")
df <- dataG %>%
  group_by(year)%>%
  summarise(pm2p5 = mean(pm2p5_lag0))
ggplot(dataG) + 
  geom_line(aes(x=date, y = pm2p5_lag0))
modlist <- readRDS("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/model_out_summary_list_weekly_adm2_dlnm_none_bivariable_8_2019_templag0humlag3.rds")
 modlist[[1]][[2]]$data$cov2
mod <- modlist[[1]][[2]]
idx <- which(mod$cp$predvar == 50)
# sum across lag columns (assumes columns are lag0..lag8)
L <- sum(mod$cp$matfit[idx, ])   # cumulative log-RR
RR <- exp(L)
AF <- (RR - 1) / RR

#### calculate number of cases above threshold of 40ug/m3 per province
dataG <- subset(data, data$year==2019)
target_munic <- "City of Johannesburg"
target_prov <- "ZAF.3_1"
obs <- dataG[dataG$GID_1 == target_prov, ]
obs <- obs %>%
  group_by(date) %>%
  summarise(pm2p5 = mean(pm2p5_lag3, na.rm=T),
            disease = sum(disease))
# obs <- dataG
### numbers from model
predvar <- mod$cp$predvar
matfit <- mod$cp$matfit
logRR <- matfit[,4]
# interpolate log-RR so we can evaluate at observed PM
Lfun <- approxfun(predvar, logRR, rule = 2)
# AF for any observed PM
AF <- function(x) {
  RR <- exp(Lfun(x))
  (RR - 1) / RR
}

# national weekly AF and attributable cases
obs$AF  = AF(obs$pm2p5)
obs$attr_cases = obs$disease * obs$AF
total_cases  = sum(obs$disease)
total_attr   = sum(obs$attr_cases)
national_PAF = total_attr / total_cases
cat(national_PAF, total_attr)

##### plot all together
pm_scaled <- with(obs, 
                  (pm2p5 - min(pm2p5)) / (max(pm2p5) - min(pm2p5)) *
                  (max(disease) - min(disease)) + min(disease)
)
scale_factor <- max(obs$disease)/max(obs$pm2p5)
pm25attr <- ggplot(obs) +
  geom_line(aes(x=date, y = pm2p5 * scale_factor, color = "PM2.5", linetype = "PM2.5")) +
  geom_line(aes(x=date, y = disease, color = "IPD cases", linetype = "IPD cases")) +
  geom_line(aes(x=date, y = attr_cases, color = "Attributable cases", linetype = "Attributable cases")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / scale_factor,
                        name = "PM2.5 (µg/m³) 3-week lag") ) +
  ylab("IPD Cases") +
  xlab("Date") +
  theme_bw()+
    scale_color_manual(name = "",values = c("IPD cases" = "black","Attributable cases" = "black","PM2.5" = "red")) +
    scale_linetype_manual(name = "",
    values = c("IPD cases" = "solid","Attributable cases" = "dashed","PM2.5" = "solid")) +
    guides(color = guide_legend(override.aes = list(linetype = c("dashed","solid","solid"))),linetype = "none") +
  theme_minimal() +
  theme(
    axis.title.y.left   = element_text(color = "black"),
    axis.title.y.right  = element_text(color = "black"),
    legend.position     = "bottom",
    legend.box          = "horizontal"
  )


#### if we got the average down to 15ug/m3 these are the number of cases avoided 
### AF with counterfactual
AF_cf <- function(x) {
  RR <- exp(Lfun(x))
  RR_cf <- exp(Lfun(40))
  (RR - RR_cf) / RR
}

# counterfactual target
PM_cf <- 15
# RR at observed and counterfactual
RR_obs <- exp(Lfun(obs$pm2p5))
RR_cf  <- exp(Lfun(PM_cf))
# AF relative to counterfactual
obs$AF_target <- (RR_obs - RR_cf) / RR_obs

# Attributable cases relative to target
obs$attr_cases_target <- obs$disease * obs$AF_target

# National summary
total_cases <- sum(obs$disease)
total_attr_cases_target <- sum(obs$attr_cases_target)
PAF_target <- total_attr_cases_target / total_cases

cat(sprintf("If PM2.5 reduced to <= %g ug/m3, %.1f cases (~%.2f%%) could be prevented\n",
            PM_cf, total_attr_cases_target, 100*PAF_target))

```

## FIGURE S3: SPEARMAN CORRELATION ACROSS LAGS 
```{r}
data <- fread(file="/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm1_weekly_lag.csv")

# Make sure your data is a data.table
setDT(data)

# Define the number of lags
lags <- 0:12
# grep("age",colnames(data), value = TRUE)
# grep("GPSC",colnames(data), value = TRUE)
# grep("Sero",colnames(data), value = TRUE)

## manual labels
color_labels = c("Eastern Cape" = "#008080", "Free State" = "#D75F00", "Gauteng" = "#967bb6" , "KwaZulu-Natal"= "#FFCCCB",
                     "Limpopo" = "#228B22", "Mpumalanga" = "#FFC000", "North West" = "#8B5A2B", "Northern Cape" = "#606060", "Western Cape" = "#1A2D6D" )
## loop through a number of covariates
cov_names_vec <- c("pm2p5","pm10","so2","o3", "hurs","absh","tas","tasmin","tasmax", "prlrmean",  "sfcWind")
cov_names_labels <- c("PM2.5","PM10","SO2","O3", "Relative Humidity (%)","Absolute Humidity (g/m3)","Mean Temp. (C)","Min. Temp. (C)","Max. Temp. (C)", "Mean Precipitation (mm)",  "Wind Speed (km/hr)")
p_list <- list()
for(c in 1:length(cov_names_vec)){
    # Create a long-format data.table for correlation results
    corr_results <- rbindlist(lapply(unique(data$adm1_name), function(prov) {
      prov_data <- data[adm1_name == prov]
      
      # For each lag, compute the Spearman correlation
      lapply(lags, function(lag) {
        lag_col <- paste0(cov_names_vec[c],"_lag", lag)
        if (lag_col %in% names(prov_data)) {
          cor_val <- cor(prov_data[[lag_col]], prov_data$disease, method = "spearman", use = "complete.obs")
        } else {
          cor_val <- NA
        }
        data.table(province = prov, lag = lag, spearman_corr = cor_val)
      }) %>% rbindlist()
    }))
    
    # Plot
    if(c==length(cov_names_vec)){
      p <- ggplot(corr_results, aes(x = lag, y = spearman_corr, color = province), alpha=0.8) +
              geom_hline(yintercept = 0, linetype = "dashed", color = "red", alpha = 0.7)+
              geom_line() +
              theme_minimal() +
              scale_color_manual(values = color_labels) +
              labs(title = cov_names_labels[c],x = "Lag (weeks)",y = "Spearman Correlation",color = "Province")  +
              theme(legend.position = "none")
    }else{
    p <- ggplot(corr_results, aes(x = lag, y = spearman_corr, color = province), alpha=0.8) +
              geom_hline(yintercept = 0, linetype = "dashed", color = "red", alpha = 0.7)+
              geom_line() +
              theme_minimal() +
              scale_color_manual(values = color_labels) +
              labs(title = cov_names_labels[c],x = "Lag (weeks)",y = "Spearman Correlation",color = "Province") +
              theme(legend.position = "none")
    }
    
    p_list[[c]] <- p
}

p <- ggplot(corr_results, aes(x = lag, y = spearman_corr, color = province), alpha=0.8) +
              geom_hline(yintercept = 0, linetype = "dashed", color = "red", alpha = 0.7)+
              geom_line() +
              theme_minimal() +
              scale_color_manual(values = color_labels) +
              labs(title = cov_names_vec[c],x = "Lag (weeks)",y = "Spearman Correlation",color = "Province") +
              theme(legend.position = "right")
pdf("/home/sbelman/Documents/env_sa_manuscript/figures/supplement/envvariablespearmanlagcorrelation.pdf", width=10, height =10)
grid.arrange(grobs = p_list, nrow=3)
dev.off()

g <- arrangeGrob(grobs = p_list, nrow = 3)  # Create grob object
ggsave("/home/sbelman/Documents/env_sa_manuscript/figures/supplement/envvariablespearmanlagcorrelation.png",
       g, width = 10, height = 10)



pdf("/home/sbelman/Documents/env_sa_manuscript/figures/supplement/envvariablespearmanlagcorrelation_LEGEND.pdf", width=3, height =3)
print(p)
dev.off()
```
## FIGURE S4: ENVIRONMENTAL DATA SUMMARY PLOTS
### days with particulate matter of 100ug/m3
### AND THRESHOLD EXCEEDANCE CALCULATIONS
```{r}
### adm1
data_unscaled <- fread("/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm2_weekly_lag.csv")
# Filter for high PM2.5 weeks and count per province and year
weeks_over_50_by_dist_year <- data_unscaled %>%
  filter(pm2p5_lag0 >= 100) %>%
  group_by(district, NAME_1, year) %>%
  summarise(weeks_over_50 = n(), .groups = "drop") %>%
  arrange(NAME_1, district,year) 
# View results
color_labels = c("Eastern_Cape" = "#008080", "Free_State" = "#D75F00", "Gauteng" = "#967bb6" , "KwaZulu-Natal"= "#FFCCCB",
                     "Limpopo" = "#228B22", "Mpumalanga" = "#FFC000", "North_West" = "#8B5A2B", "Northern_Cape" = "#606060", "Western_Cape" = "#1A2D6D" )
weeksover <- ggplot(weeks_over_50_by_dist_year, aes(x = year, y = weeks_over_50, color = NAME_1, group = district)) +
  geom_line(size = 1, alpha = 0.7) +
  # geom_point(alpha = 0.7) +
  scale_color_manual(values=color_labels)+
  labs(x = "Year",
     y = expression("N Weeks " * PM[2.5] * " > 40 (" * µ * "g/m"^3 * ")"),
    color = "Province"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )+
  theme(legend.position = "none")

weeks_over_50_by_dist_year[which(weeks_over_50_by_dist_year$weeks_over_50>10),]
####### tasmax vs relative humidity
data_unscaled <- fread("/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm1_weekly_lag.csv")
color_labels = c("Eastern Cape" = "#008080", "Free State" = "#D75F00", "Gauteng" = "#967bb6" , "KwaZulu-Natal"= "#FFCCCB",
                     "Limpopo" = "#228B22", "Mpumalanga" = "#FFC000", "North West" = "#8B5A2B", "Northern Cape" = "#606060", "Western Cape" = "#1A2D6D" )

tmpvhum <- ggplot(data_unscaled, aes(x = tasmax_lag0, y = hurs_lag0, group=province,color=province)) +
  geom_point(alpha = 0.4) +
  scale_color_manual(values=color_labels)+
  labs(
    x = "Maximum Temperature (°C)",
    y = "Relative Humidity (%)"
  ) +
  theme_minimal()+
  theme(legend.position = "right")
pm2p5vhum <- ggplot(data_unscaled, aes(x = pm2p5_lag0, y = hurs_lag0, group=province,color=province)) +
  geom_point(alpha = 0.4) +
  scale_color_manual(values=color_labels)+
  labs(
    x = "PM2.5",
    y = "Relative Humidity (%)"
  ) +
  theme_minimal()+
  theme(legend.position = "none")

pm2p5vso2 <- ggplot(data_unscaled, aes(x = pm2p5_lag0, y = so2_lag0, group=province,color=province)) +
  geom_point(alpha = 0.2) +
  scale_color_manual(values=color_labels)+
  labs(
    x = "PM2.5",
    y = "SO2"
  ) +
  theme_minimal()+
  theme(legend.position = "right")

prlrvhum <- ggplot(data_unscaled, aes(x = prlrsum_lag0, y = hurs_lag0, group=province,color=province)) +
  geom_point(alpha = 0.4) +
  scale_color_manual(values=color_labels)+
  labs(
    x = "Precipitation (mm)",
    y = "Relative Humidity (%)"
  ) +
  theme_minimal()+
  theme(legend.position = "none")

##### pm2.5 and pm10
hist_pm25 <- ggplot(data_unscaled, aes(x = pm2p5_lag0, group=province, fill=province)) +
  geom_histogram( color = "black", bins = 30) +
    scale_fill_manual(values=color_labels)+
  labs(x = expression("PM"[2.5]*" (µg/m³)"),
    y = "Frequency"
  ) +
  theme_minimal() 

pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/supplement/envtmpvhum_descriptions.pdf", width = 4,height=4)
tmpvhum
dev.off()
pdf(file="/home/sbelman/Documents/env_sa_manuscript/figures/supplement/env_descriptions.pdf", width = 13,height=4)
(weeksover + hist_pm25) + 
  patchwork::plot_annotation(tag_levels = 'a')
dev.off()
(weeksover + hist_pm25) + 
  patchwork::plot_annotation(tag_levels = 'a')
ggsave(file="/home/sbelman/Documents/env_sa_manuscript/figures/supplement/env_descriptions.png", width = 10,height=4)

########## THRESHOLD EXCEEDANCE ################################################
# Count weeks ≥100 for each district-year
weeks_over_100 <- data_unscaled %>%
  group_by(district, NAME_1, year) %>%
  summarise(weeks_over_100 = sum(pm2p5_lag0 >= 50), .groups = "drop")

# Keep only district-years that had ≥1 exceedance
districts_with_exceedance <- weeks_over_100 %>%
  filter(weeks_over_100 > 0)

# For each year: 
#   - N of districts that exceeded
#   - Avg # of exceedance weeks (among only those districts)
summary_by_year <- districts_with_exceedance %>%
  group_by(year) %>%
  summarise(
    districts_exceeding = n(),
    avg_weeks_exceeding = mean(weeks_over_100),
    .groups = "drop"
  )

summary_by_year
mean(summary_by_year[which(summary_by_year$year<=2010),]$avg_weeks_exceeding,na.rm=T)
mean(summary_by_year[which(summary_by_year$year>2010&summary_by_year$year<=2923),]$avg_weeks_exceeding,na.rm=T)
mean(summary_by_year[which(summary_by_year$year>2015&summary_by_year$year<=2023),]$avg_weeks_exceeding,na.rm=T)
```


## FIGURE S8: SENSITIVITY ACROSS CONFOUNDERS IN AIR POLLUTION MODELS
```{r}
##### compare CBpm2p5 + CBtas + absh model TO CBpm2p5 + tas + absh model
dfth1 <- fread("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/nointeraction_results_fits_weekly_adm2_bivariable_8_2019.csv")
dfth <- subset(dfth , dfth$covariate == "pm2p5_lag0" & dfth$cov2 == "absh_lag3")
dfthC <- subset(dfth , dfth$lag_num == 3)
dfthC$cumrr <- exp(dfthC$cumulative_fit)
dfthC$cumlowerCI_rr <- exp(dfthC$cum_lowerCI)
dfthC$cumupperCI_rr <- exp(dfthC$cum_upperCI)

#### temperature alone
dft <- subset(dfth1 , dfth1$covariate == "pm2p5_lag0" & dfth1$cov2 == "tas_lag3")
dftC <- subset(dft , dft$lag_num == 3)
dftC$cumrr <- exp(dftC$cumulative_fit)
dftC$cumlowerCI_rr <- exp(dftC$cum_lowerCI)
dftC$cumupperCI_rr <- exp(dftC$cum_upperCI)
## read in double cross basis
dfCB2 <- fread("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/modifiers/nointeraction_dlnm2x_results_fits_weekly_adm2_tasabsh_8_2019.csv")
dfCB2 <- subset(dfCB2 , dfCB2$covariate == "pm2p5" )
dfCB2C <- subset(dfCB2 , dfCB2$covariate == "pm2p5" & dfCB2$lag_num == 3)
dfCB2C$cumrr <- exp(dfCB2C$cumulative_fit)
dfCB2C$cumlowerCI_rr <- exp(dfCB2C$cum_lowerCI)
dfCB2C$cumupperCI_rr <- exp(dfCB2C$cum_upperCI)

## interaction version
dfint <- fread("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/modifiers/envmod_results_fits_weekly_adm2_allmodifiers_propprov_8_2019.csv")
dfint <- subset(dfint , dfint$covariate == "pm2p5_lag0" &  dfint$interaction_level =="med")
dfintC <- subset(dfint , dfint$lag_num == 3 )
dfintC$cumrr <- exp(dfintC$cumulative_fit)
dfintC$cumlowerCI_rr <- exp(dfintC$cum_lowerCI)
dfintC$cumupperCI_rr <- exp(dfintC$cum_upperCI)

## no covariates
dfnone <- fread("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/univariable/nointeraction_results_fits_weekly_adm2_8_2019.csv")
dfnone <- subset(dfnone, dfnone$covariate=="pm2p5_lag0"& dfnone$lag == "lag3")
dfnone$cumrr <- exp(dfnone$cumulative_fit)
dfnone$cumlowerCI_rr <- exp(dfnone$cum_lowerCI)
dfnone$cumupperCI_rr <- exp(dfnone$cum_upperCI)

ggplot() +
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_line(data = dfthC, aes(x=predvar, y = cumrr), color = "darkgreen") +
  geom_ribbon(data=dfthC, aes(x=predvar, ymin = cumlowerCI_rr, ymax = cumupperCI_rr), fill = "darkgreen", alpha = 0.2) +
  geom_line(data = dfCB2C, aes(x=predvar, y = cumrr), color = "darkblue") +
  geom_ribbon(data=dfCB2C, aes(x=predvar, ymin = cumlowerCI_rr, ymax = cumupperCI_rr), fill = "darkblue", alpha = 0.2) +
  theme_bw()+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10", limits = c(0.8,1.9), breaks = c(0.8, 1, 1.8))+
  # facet_wrap(.~lag_week, nrow = 1)+
  xlab(expression(paste('Concentration (', mu, 'g/m'^3, ')')))+  
  geom_line(data = dfnone, aes(x=predvar, y = cumrr), color = "black", linetype = "dashed") +
  geom_ribbon(data=dfnone, aes(x=predvar, ymin = cumlowerCI_rr, ymax = cumupperCI_rr), fill = "black",linetype = "dashed", alpha = 0.2) +
  theme_bw()+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "right",legend.title = element_blank())



dftC$type <- dfnone$type <- dfthC$type <- dfCB2C$type <- dfintC$type <- NULL
dfthC$type <- "CBpm2.5 + tas_lag3+absh_lag3"
dfCB2C$type <- "CBpm2.5 + CBtas + absh_lag3"
dfintC$type <- "CBpm2.5*tas_lag3 + absh_lag3"
dftC$type <- "CBpm2.5 + CBtas"
dfnone$type <- "CBpm2.5"


rbind(
  dfnone[which(dfnone$predvar==50),c("type","cumrr","cumlowerCI_rr","cumupperCI_rr")],
  dftC[which(dftC$predvar==50),c("type","cumrr","cumlowerCI_rr","cumupperCI_rr")],
dfthC[which(dfthC$predvar==50),c("type","cumrr","cumlowerCI_rr","cumupperCI_rr")],
dfCB2C[which(dfthC$predvar==50),c("type","cumrr","cumlowerCI_rr","cumupperCI_rr")],
dfintC[which(dfthC$predvar==50),c("type","cumrr","cumlowerCI_rr","cumupperCI_rr")])
```

#### Figure S14: visualize count of sequences for Figure S14
```{r}
data2<- fread(file="/home/sbelman/Documents/BRD/SouthAfrica/disease/SA_disease_point_base.csv",quote=FALSE, header = TRUE)
out <- data2 %>%
  group_by(province,year) %>%
  summarise(count_seq = sum(sequenced, na.rm = T))

color_labels = c("Eastern Cape" = "#008080", "Free State" = "#D75F00", "Gauteng" = "#967bb6" , "KwaZulu-Natal"= "#FFCCCB",
                     "Limpopo" = "#228B22", "Mpumalanga" = "#FFC000", "North West" = "#8B5A2B", "Northern Cape" = "#606060", "Western Cape" = "#1A2D6D" )
countpyeryear <- ggplot(out[which(out$year<2020),]) + 
  geom_line(aes(x= year, y = count_seq, group = province, color = province))+
  theme_bw() +
  theme(element_text(size = 15))+
  scale_color_manual(values = color_labels)+
  xlab("Year")+
  ylab("Count") +
  labs(color = "Province")
pdf("/home/sbelman/Documents/env_sa_manuscript/figures/supplement/seqcountperyear.pdf", width = 8, height = 4)
print(countpyeryear)
dev.off()
```

#### Figure S7 and S12 VISUALIZE SURFACE PLOTS FOR SUPPLEMENT Figure S12 and Figure S7
```{r}
library(GHRmodel)

### air quality surface plots
#PM2.5 surface plot with absolute humidity and temperature controls
mods <- readRDS("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/model_out_summary_list_weekly_adm2_dlnm_none_bivariable_8_2019_templag0humlag3.rds")

## main model surface plot
mods1 <- mods[[1]][[2]]
pmabshtmp <- plot_coef_crosspred(
  crosspred = mods1$cp,  # Crosspred object containing model predictions
  type = "heatmap",           # Create a 2D heatmap of effects over exposure and lag values
  exp = TRUE,
  heat_lims = c(0.78,1.21),
  palette = "-RdBu",
  xlab = expression("PM"[2.5]*" (µg/m³)"))

### plot all air pollutions in a list
ap_plot_list <- list()
pollutant_names <- c("PM2.5", "PM10", "SO2")
pollutant_names <- c("PM₂.₅", "PM₁₀", "SO₂")

counter <- 1
for (p in seq_along(mods)) {
  heat_lim_vec <- list(c(0.94,1.05),c(0.94,1.05),c(0.9,1.09))
  for (i in seq_along(mods[[p]])) {
    cov1 <- mods[[p]][[i]]$data$cov
    cov2 <- mods[[p]][[i]]$data$cov2
    if(cov2%in%c("hurs_lag3","absh_lag3")){
      ap_plot_list[[counter]] <- plot_coef_crosspred(
      crosspred = mods[[p]][[i]]$cp,
      type = "heatmap",
      heat_lims = heat_lim_vec[[p]],
      exp = TRUE,
      palette = "-RdBu",
      xlab = "Lag Week",
      ylab = paste0(pollutant_names[p]," µg/m³ (",gsub("_lag3","",cov2), " + tas)"))
    }else{
      ap_plot_list[[counter]] <- plot_coef_crosspred(
      crosspred = mods[[p]][[i]]$cp,
      type = "heatmap",
      heat_lims = heat_lim_vec[[p]],
      exp = TRUE,
      palette = "-RdBu",
      xlab = "Lag Week",
      ylab = paste0(pollutant_names[p]," µg/m³ (",gsub("_lag0","",cov2), ")"))
    }
    
    
    counter <- counter + 1
  }
}


pdf("/home/sbelman/Documents/env_sa_manuscript/figures/supplement/airpollution_surfaces.pdf", width = 12, height = 8)
wrap_plots(ap_plot_list)
dev.off()

wrap_plots(ap_plot_list)
ggsave(file = "/home/sbelman/Documents/env_sa_manuscript/figures/supplement/airpollution_surfaces.png", width = 12, height = 8)
dev.off()

## meteorlogical crosspred
modsmet <- readRDS("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/univariable/model_out_summary_list_weekly_adm2_dlnm_none_8_2019_meteo.rds")
## wrap plots from plot_coef_crosspred
color_vec <- c(rep("-RdBu",3),rep("Green-Brown",3),"-RdBu",rep("Green-Brown",4))
metplot_list <- list()
for(i in 1:length(modsmet)){
  metplot_list[[i]] <- plot_coef_crosspred(
  crosspred = modsmet[[i]]$cp,  # Crosspred object containing model predictions
  type = "heatmap",           # Create a 2D heatmap of effects over exposure and lag values
  exp = TRUE,                 # Exponentiate effects to show relative risks
  palette = "-RdBu",         
  xlab = "Lag Week",
  ylab = gsub("_lag0","",modsmet[[i]]$data$cov))
}

pdf("/home/sbelman/Documents/env_sa_manuscript/figures/supplement/meteorological_surfaces.pdf", width = 12, height = 6)
wrap_plots(metplot_list)
dev.off()

wrap_plots(metplot_list)
ggsave(file = "/home/sbelman/Documents/env_sa_manuscript/figures/supplement/meteorological_surfaces.png", width = 12, height = 6)
dev.off()
```

## Figures S9-S11:  SUPPLEMENTAL FIGURES GAUTENG 
# Gauteng Alone Compared to South Africa Sensitivity (using fit_list South Africa from section starting at line 950)
```{r}
#Combine Univariable results for each for meteorological covariates and then only include the South Africa model with tas+absh for PM
time = "weekly"
space = "adm2"
max_lag = 8
prov_name = "Gauteng"
## read in Gauteng
  ## meteorological only 
nointgaut <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/suboutcomes/province/nointeraction_results_fits_",time,"_",space,"_",max_lag,"week_2019_",prov_name,".csv"))
nointgaut$cov2 <- "none"
nointgaut <- subset(nointgaut, nointgaut$covariate%notin%c("pm2p5_lag0","pm10_lag0","so2_lag0"))
nointgaut$type <- "Gauteng"
## pollution only
biv_nointgaut <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/suboutcomes/province/nointeraction_results_fits_",time,"_",space,"_bivariable_8week_2019_",prov_name,"_temp0hum3.csv"))
biv_nointgaut <- subset(biv_nointgaut, biv_nointgaut$covariate%in%c("pm2p5_lag0","pm10_lag0","so2_lag0") & biv_nointgaut$cov2=="absh_lag3")
biv_nointgaut$type <- "Gauteng"
## pollution only NO GAUTENG
biv_nointnogaut <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/suboutcomes/province/nointeraction_results_fits_",time,"_",space,"_bivariable_8week_2019_exclude",prov_name,"_temp0hum3.csv"))
biv_nointnogaut <- subset(biv_nointnogaut, biv_nointnogaut$covariate%in%c("pm2p5_lag0","pm10_lag0","so2_lag0") & biv_nointnogaut$cov2=="absh_lag3")
biv_nointnogaut$type <- "NoGauteng"
fit_list_gaut <- rbind(biv_nointgaut,nointgaut,biv_nointnogaut)
# fit_list_gaut <- rbind(biv_nointgaut,nointgaut)

fit_list_gaut$lag_week <- paste0("Week",fit_list_gaut$lag_num)
fit_list_gaut$cov <- gsub("_lag0","",fit_list_gaut$covariate)
fit_list_gaut$lag_week <- factor(fit_list_gaut$lag_week, levels = paste0("Week",seq(0,8,1)), labels = paste0("Week",seq(0,8,1)))
# convert to relative risks
fit_list_gaut$rr <- exp(fit_list_gaut$fit)
fit_list_gaut$lowerCI_rr <- exp(fit_list_gaut$lowerCI)
fit_list_gaut$upperCI_rr <- exp(fit_list_gaut$upperCI)
fit_list_gaut$cumrr <- exp(fit_list_gaut$cumulative_fit)
fit_list_gaut$cumlowerCI_rr <- exp(fit_list_gaut$cum_lowerCI)
fit_list_gaut$cumupperCI_rr <- exp(fit_list_gaut$cum_upperCI)
# fit_list_gaut$type <- "Gauteng"
cols <- colnames(fit_list_gaut)[which(colnames(fit_list_gaut)!="type")]

## change name of south africa one to mitigate confusion
fit_list_SA <- fit_list
fit_list_SA <- subset(fit_list_SA, fit_list_SA$space_time==paste0("weekly_",space))
fit_list_SA <- fit_list_SA[,..cols]
fit_list_SA$type <- "South Africa"
fit_list_SA[which(fit_list_SA$predvar==50&fit_list_SA$cov2=="absh_lag3 + tas_lag0" & fit_list_SA$cov=="pm2p5")]

fit_listcomp <- rbind(fit_list_gaut, fit_list_SA)
fit_listcomp <- subset(fit_listcomp, fit_listcomp$cov2%in%c("absh_lag3", "absh_lag3 + tas_lag0","none"))

## drop any air pollutants with no cov2
fit_listcomp <- fit_listcomp %>%
  filter(!(cov %in% c("pm2p5","pm10","so2") & cov2 == "none"))
fit_listcomp$cov2[which(fit_listcomp$cov2 == "absh_lag3")] <- "absh_lag3 + tas_lag0"

########### cumulative plot of all on grid ####
fit_listcompcum <- subset(fit_listcomp, fit_listcomp$lag=="lag3")
# turn to factors and label so that metrics are included
fit_listcompcum$covfact <- factor(fit_listcompcum$cov, levels = c("tas", "tasmin", "tasmax",
                                                                  "prlrmean","prlrsum","spei3","spei6",
                                                                  "absh","hurs", "sfcWind", "o3",
                                                                 "pm2p5","pm10","so2"),
                              labels = c("Temp (°C) ", "Min Temp (°C)", "Max Temp (°C)",
                                                                  "Mean Precip (mm)","Total Precip (mm)","SPEI 3-month","SPEI 6-month",
                                                                  "Absolute Humidity (g/m3)","Relative Humidity (%)", "Wind Speed (km/hr)", "Ozone (µg/m³)",
                                                                 "PM2.5 (µg/m³)","PM10 (µg/m³)","SO2 (µg/m³)"))



################## PLOTS FOR GAUTENG ALONE AND SOUTH AFRICA OVERALL ############
cumplots <- ggplot(fit_listcompcum[which(fit_listcompcum$type!="NoGauteng")])+
  geom_line(aes(x=var,y=cumrr,group=interaction(type,covfact), color = type),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(type,covfact), fill=type),alpha=0.4)+
  scale_color_manual(values = c("Gauteng" = "purple","", "South Africa" = "black"))+
  scale_fill_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10")+
  # scale_y_continuous(trans="log10", limits = c(0.85,1.15), breaks = c(0.85, 1, 1.15))+
  labs(color = "", fill = "")+
  facet_wrap(.~covfact, scales = "free")+
  xlab("Exposure Value")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), strip.text = element_text(size=11), axis.text.x = element_text(angle = 45,hjust=1, size=13))
pdf("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/gautengsouthAfrica_cumulative.pdf", width = 11, height = 8)
print(cumplots)
dev.off()
print(cumplots)
ggsave("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/gautengsouthAfrica_cumulative.png", width = 11, height = 8)
dev.off()



########### lagged plot of all meteorological on grid ####
# turn to factors and label so that metrics are included
fit_listcomp$covfact <- factor(fit_listcomp$cov, levels = c("tas", "tasmin", "tasmax",
                                                                  "prlrmean","prlrsum","spei3","spei6",
                                                                  "absh","hurs", "sfcWind", "o3",
                                                                 "pm2p5","pm10","so2"),
                              labels = c("Temp (°C) ", "Min Temp (°C)", "Max Temp (°C)","Mean Precip (mm)","Total Precip (mm)","SPEI 3-month","SPEI 6-month","Absolute Humidity (g/m3)","Relative Humidity (%)", "Wind Speed (km/hr)", "Ozone (µg/m³)","PM2.5 (µg/m³)","PM10 (µg/m³)","SO2 (µg/m³)"))

## temperature
templaggautplot <- ggplot(fit_listcomp[which(fit_listcomp$cov %in% c("tas","tasmin","tasmax"))])+
  geom_line(aes(x=var,y=rr,group=interaction(type,covfact,lag_num), color = type),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(type,covfact,lag_num), fill=type),alpha=0.4)+
  scale_color_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  scale_fill_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.7,1.3), breaks = c(0.7, 1, 1.3))+
  labs(color = "", fill = "")+
  xlab("Exposure Value")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13)
        ,strip.text = element_blank(),strip.background = element_blank()
  )+
    facet_wrap(covfact~lag_num,scales = "free_x", ncol = 9, labeller = labeller(lag_num = function(x) ""))

 templaggautplot
 
 
 
 ## humidity, wind, and ozone
humlaggautplot <- ggplot(fit_listcomp[which(fit_listcomp$cov %in% c("absh","hurs","sfcWind","o3"))])+
  geom_line(aes(x=var,y=rr,group=interaction(type,covfact,lag_num), color = type),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(type,covfact,lag_num), fill=type),alpha=0.4)+
  scale_color_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  scale_fill_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.8,1.2), breaks = c(0.8, 1, 1.2))+
  labs(color = "", fill = "")+
  xlab("Exposure Value")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13)
        ,strip.text = element_blank(),strip.background = element_blank()
  )+
    facet_wrap(covfact~lag_num,scales = "free", ncol = 9, labeller = labeller(lag_num = function(x) ""))

 humlaggautplot
 
 
 
 ## precipitaion 
prlrlaggautplot <- ggplot(fit_listcomp[which(fit_listcomp$cov %in% c("prlrmean","prlrsum"))])+
  geom_line(aes(x=var,y=rr,group=interaction(type,covfact,lag_num), color = type),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(type,covfact,lag_num), fill=type),alpha=0.4)+
  scale_color_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  scale_fill_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.4,1.8), breaks = c(0.4, 1, 1.8))+
  labs(color = "", fill = "")+
  xlab("Exposure Value")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13)
        ,strip.text = element_blank(),strip.background = element_blank()
  )+
    facet_wrap(covfact~lag_num,scales = "free", ncol = 9, labeller = labeller(lag_num = function(x) ""))
 ## SPEI
speilaggautplot <- ggplot(fit_listcomp[which(fit_listcomp$cov %in% c("spei3","spei6"))])+
  geom_line(aes(x=var,y=rr,group=interaction(type,covfact,lag_num), color = type),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(type,covfact,lag_num), fill=type),alpha=0.4)+
  scale_color_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  scale_fill_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.85,1.2), breaks = c(0.85, 1, 1.2))+
  labs(color = "", fill = "")+
  xlab("Exposure Value")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13)
        ,strip.text = element_blank(),strip.background = element_blank()
  )+
    facet_wrap(covfact~lag_num,scales = "free", ncol = 9, labeller = labeller(lag_num = function(x) ""))

 

 ## POLLUTANTS
aqlaggautplot <- ggplot(fit_listcomp[which(fit_listcomp$cov %in% c("pm2p5","pm10","so2"))])+
  geom_line(aes(x=var,y=rr,group=interaction(type,covfact,lag_num), color = type),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(type,covfact,lag_num), fill=type),alpha=0.4)+
  scale_color_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  scale_fill_manual(values = c("Gauteng" = "purple", "South Africa" = "black"))+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  # scale_y_continuous(trans="log10")+
  scale_y_continuous(trans="log10", limits = c(0.85,1.2), breaks = c(0.85, 1, 1.2))+
  labs(color = "", fill = "")+
  xlab("Exposure Value")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13)
        ,strip.text = element_blank(),strip.background = element_blank()
  )+
    facet_wrap(covfact~lag_num,scales = "free", ncol = 9, labeller = labeller(lag_num = function(x) ""))

pdf("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/gautengsouthAfrica_temp_weekly.pdf", width = 13, height = 4.5)
print(templaggautplot)
dev.off() 

pdf("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/gautengsouthAfrica_hum_weekly.pdf", width = 13, height = 6)
print(humlaggautplot)
dev.off() 

pdf("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/gautengsouthAfrica_prlr_weekly.pdf", width = 13, height = 3)
print(prlrlaggautplot)
dev.off() 
 
pdf("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/gautengsouthAfrica_spei_weekly.pdf", width = 13, height = 3)
print(speilaggautplot)
dev.off() 

pdf("/home/sbelman/Documents/env_sa_manuscript/figures/figure3/gautengsouthAfrica_aq_weekly.pdf", width = 13, height = 4.5)
print(aqlaggautplot)
dev.off() 
```
##### FIGURE S9 and S10
###### GAUTENG VS SOUTH AFRICA
```{r}
############# CUMULATIVE EFFECT PLOTS ############
max_lag <- 8
prov_name = "Gauteng"
space = "adm2"
endyear = 2019
time = "weekly"
noint <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/suboutcomes/province/nointeraction_results_fits_",time,"_",space,"_",max_lag,"week_",endyear,"_",prov_name,".csv"))
noint$cov <- gsub("_lag0","",noint$covariate)
noint$prov <- "Gauteng"

## WHOLE COUNTRY
m1 <- fread(file = paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/univariable/nointeraction_results_fits_weekly_adm2_8_2019.csv"))
m1$cov <- gsub("_lag0","",m1$covariate)
m1$prov <- "South_Africa"
# bind
nointall <- rbind(noint,m1)
cumgaut <- ggplot(nointall[which(nointall$cov%notin%c("spi3","spi6")),])+
  geom_hline(yintercept=1, linetype="dashed",color="red")+
  geom_line(aes(x = var, y = exp(cumulative_fit), group=interaction(cov,prov),color=prov))+
  geom_ribbon(aes(x = var, ymin = exp(cum_lowerCI), ymax= exp(cum_upperCI), group=interaction(cov,prov), fill=prov), alpha=0.5)+
  theme_bw()+
  scale_y_continuous(trans = "log10")+
  # scale_color_manual(values = color_labels2)+
  # scale_fill_manual(values = color_labels2)+
  # scale_linetype_manual(values = linetype_labels2) +
  theme(legend.title = element_blank(), legend.text = element_text(size=13))+
  theme(axis.text = element_text(size = 13), axis.title = element_text(size =13), strip.text = element_text(size=13), axis.text.x = element_text(angle = 45, hjust =1) )+
  ylab("Relative Risk")+
  # ylim(0.1,3)+
  # xlab("Variable")+
  facet_wrap(cov~., scales="free")
pdf("/home/sbelman/Documents/env_sa_manuscript/figures/gauteng_vs_southafrica_cumulative.pdf",width=10,height=8)
print(cumgaut)
dev.off()
print(cumgaut)
ggsave("/home/sbelman/Documents/env_sa_manuscript/figures/gauteng_vs_southafrica_cumulative.png",width=10,height=8)
dev.off()


############# LAG EFFECT Plots ############
color_labels2 = c("Gauteng" = "purple", "South_Africa" = "black")
linetype_labels2 = c("Gauteng" = "dashed", "South_Africa" = "solid")
nointalltmp <- subset(nointall, nointall$cov%in%c("pm2p5","pm10","so2","hurs","tasmax","prlrsum") & nointall$lag_num<9)
nointalltmp$cov <- factor(nointalltmp$cov, levels = c("hurs","prlrsum","tasmax","pm2p5","pm10","so2"))
laggaut <- ggplot(nointalltmp)+
  geom_hline(yintercept=1, linetype="dashed",color="red")+
  geom_line(aes(x = var, y = exp(fit), group=interaction(cov,prov, lag_num),color=prov, linetype = prov))+
  geom_ribbon(aes(x = var, ymin = exp(lowerCI), ymax= exp(upperCI), group=interaction(cov,prov,lag_num), fill=prov), alpha=0.5)+
  theme_bw()+
  ylab("Relative Risk")+
  scale_y_continuous(trans = "log10")+
  scale_color_manual(values = color_labels2)+
  scale_fill_manual(values = color_labels2)+
  scale_linetype_manual(values = linetype_labels2) +
  # ylim(0.1,3)+
  xlab("Variable")+
  theme(legend.title = element_blank(), legend.text = element_text(size=13))+
  theme(axis.text = element_text(size = 13), axis.title = element_text(size =13), strip.text = element_text(size=13), axis.text.x = element_text(angle = 45, hjust =1) )+
  facet_wrap(cov~lag, scales="free", ncol=9)

### save lags
print(laggaut)
ggsave("/home/sbelman/Documents/env_sa_manuscript/figures/gauteng_vs_southafrica_lags.png",width=20,height=15)
dev.off()

pdf("/home/sbelman/Documents/env_sa_manuscript/figures/gauteng_vs_southafrica_lags.pdf",width=20,height=15)
print(laggaut)
dev.off()
```
## Figure S11: PLOT WITH GAUTENG ALONE, NO GAUTENG, AND ALL OF SOUTH AFRICA. RELIES ON TOP PART OF CHUNK ABOVE
```{r}
fit_listcompcum$type <- factor(fit_listcompcum$type, levels = c("NoGauteng","Gauteng","South Africa"))
cumplotsno <- ggplot(fit_listcompcum[which(fit_listcompcum$cov%in%c("pm2p5","pm10","so2"))])+
  geom_line(aes(x=var,y=cumrr,group=interaction(type,covfact), color = type),alpha=0.7)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(type,covfact), fill=type),alpha=0.4)+
  scale_color_manual(values = c("Gauteng" = "purple","NoGauteng" = "darkgreen", "South Africa" = "black"))+
  scale_fill_manual(values = c("Gauteng" = "purple", "NoGauteng" = "darkgreen", "South Africa" = "black"))+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10")+
  # scale_y_continuous(trans="log10", limits = c(0.85,1.15), breaks = c(0.85, 1, 1.15))+
  labs(color = "", fill = "")+
  facet_grid(type~covfact, scales = "free_x")+
  xlab("Exposure Value")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), strip.text = element_text(size=11), axis.text.x = element_text(angle = 45,hjust=1, size=13))

  pdf("/home/sbelman/Documents/env_sa_manuscript/figures/supplement/NOgautengsouthAfrica_adm2_cumulative.pdf", width = 6, height = 4)
  print(cumplotsno)
  dev.off()



#### point plot of each lag and cumulative for all 3 types
fitsweeks50 <- fit_listcomp[which(fit_listcomp$cov%in%c("pm2p5")&fit_listcomp$predvar==50),
                       c("rr","lowerCI_rr","upperCI_rr","cov","covfact","type","lag_num","lag_week")]
fitscum50 <- fit_listcompcum[which(fit_listcompcum$cov%in%c("pm2p5")&fit_listcompcum$predvar==50),
                             c("cumrr","cumlowerCI_rr","cumupperCI_rr","cov","covfact","type","lag_num","lag_week")]
fitscum50$lag_num <- 10
fitscum50$lag_week <- "cumulative"
colnames(fitscum50) <-  c("rr","lowerCI_rr","upperCI_rr","cov","covfact","type","lag_num","lag_week")
fits50 <- rbind(fitsweeks50,fitscum50)
fits50$modtype <- "weekly"
fits50$modtype[which(fits50$lag_num==10)] <- "cumulative"
fits50$modtype <- factor(fits50$modtype,levels=c("weekly","cumulative"))

plotfit50 <- ggplot(fits50) +
      geom_hline(yintercept=1,color="red",linetype="dashed",alpha=0.6) +
      geom_pointrange(aes(x=lag_num,ymin=lowerCI_rr,ymax=upperCI_rr,y=rr,color=type,group=type),alpha=0.8,position=position_dodge(width=0.45)) +
      scale_color_manual(values=c("Gauteng"="purple","NoGauteng"="darkgreen","South Africa"="black")) +
      scale_y_continuous(trans="log10") +
      ylab(expression("Relative Risk at 50 "*mu*"g/m"^3*" PM2.5")) +
      scale_x_continuous(breaks=0:8,labels=paste0("week",1:9)) +
      facet_grid(.~modtype,scales="free_x",space="free_x",labeller=labeller(modtype=c("weekly"="Weekly lag","cumulative"="Cumulative"))) +
      theme_bw() +
      theme(strip.background=element_rect(fill="white"),strip.text=element_text(size = 9),axis.title.x=element_blank(),
            axis.text = element_text(size=12), axis.text.x = element_text(angle = 45, hjust =1),legend.title = element_blank())

  pdf("/home/sbelman/Documents/env_sa_manuscript/figures/supplement/NOgautengsouthAfrica_adm2_50ugm3weekly.pdf", width = 9, height = 3)
  print(plotfit50)
  dev.off()
```

#### Figure S12: plot summary of proportions of GPSCs for supplement S12
```{r}
data2<- fread(file="/home/sbelman/Documents/env_sa_manuscript/input_datasets/disease/SA_disease_point_base.csv",quote=FALSE, header = TRUE)
gpscdf <- as.data.table(table(data2$GPSC))[order(-N)]
shp <-st_read("/home/sbelman/Documents/env_sa_manuscript/input_datasets/shps/gadm41_namematch_ZAF_1.shp")


#  Filter targets: N >= 100 or specific ID "8"
target_gpscs <- gpscdf %>% 
  filter(N >= 100 | V1 == "8") %>% 
  pull(V1)

# Prepare Panel A: Time series of proportions per week
df_a <- data2 %>%
  mutate(Week = floor_date(as.Date(date), "week")) %>%
  # Step 1: Calculate total sequences per week across the whole dataset
  group_by(Week) %>%
  mutate(weekly_total_sequenced = sum(sequenced, na.rm = TRUE)) %>%
  # Step 2: Filter for the lineages we want to plot (and only sequenced rows)
  filter(sequenced == 1, GPSC %in% target_gpscs) %>%
  # Step 3: Count instances of each GPSC per week
  group_by(Week, GPSC, weekly_total_sequenced) %>%
  summarise(gpsc_count = n(), .groups = "drop") %>%
  # Step 4: Calculate the proportion
  mutate(proportion = gpsc_count / weekly_total_sequenced,
    GPSC_label = paste0("GPSC", GPSC))

# Prepare Panel B: Spatial distribution per lineage
# (Proportion of each GPSC found within each province)
df_b <- data2 %>%
  # 1. Filter to only rows that were actually sequenced
  filter(sequenced == 1) %>%
  # 2. Calculate the total sequences for each province across all lineages
  group_by(province) %>%
  mutate(province_total_sequenced = n()) %>%
  ungroup() %>%
  # 3. Now filter for only your target GPSCs
  filter(GPSC %in% target_gpscs) %>%
  # 4. Group by province and GPSC to get the numerator
  group_by(province, GPSC, province_total_sequenced) %>%
  summarise(gpsc_prov_count = n(), .groups = "drop") %>%
  # 5. Calculate the proportion based on the province total we saved earlier
  mutate(prop = gpsc_prov_count / province_total_sequenced,
    GPSC_label = paste0("GPSC", GPSC))


## reorder
label_order <- paste0("GPSC", target_gpscs)
df_a <- df_a %>%
  mutate(GPSC = factor(GPSC, levels = target_gpscs),
         GPSC_label = factor(GPSC_label, levels = label_order))

df_b <- df_b %>%
  mutate(GPSC = factor(GPSC, levels = target_gpscs),
         GPSC_label = factor(GPSC_label, levels = label_order))



# 4. Create Plots
plot_a <- ggplot(df_a, aes(x = Week, y = proportion)) +
  geom_col(width = 7, fill = "black", color = "black") + 
  facet_wrap(~GPSC_label, scales = "free", ncol = 3) +
  labs(y = "Proportion", x = "Date") +
  theme_bw()+
  theme(strip.background = element_rect(fill = "white"),axis.text.x = element_text(angle=45, hjust =1))+
  theme(element_text(size=12))

colnames(shp)[which(colnames(shp)=="NAME_1")] <- "province"
df_b$province <- gsub("_"," ",df_b$province)

shpprop <- shp %>% 
  inner_join(df_b, by = "province") 
shpprop$GPSC_label <- gsub("_count","",shpprop$GPSC_label)
shpprop$GPSC_label <- factor(shpprop$GPSC_label, levels = paste0("GPSC",target_gpscs))
plot_b <- ggplot(shpprop) +
  geom_sf(aes(fill = prop), color = "white", size = 0.1) +
  facet_wrap(~GPSC_label, ncol = 3) +
  scale_fill_viridis_c(option = "mako", name = "proportion") + 
  theme_void() +
  theme(element_text(size=12))

pdf("/home/sbelman/Documents/env_sa_manuscript/figures/figure5/gpsc_proportions_perweek.pdf",width = 6, height =6)
print(plot_a)
dev.off()
pdf("/home/sbelman/Documents/env_sa_manuscript/figures/figure5/gpsc_proportions_perprovince.pdf",width = 6, height =6)
print(plot_b)
dev.off()
```
## Figure S22: SURFACE PLOTS FOR GPSC MODELS 
```{r}
library(GHRmodel)
library(patchwork)
gpsc_vec <- c("3","17","2","14","10","5","13","26","1","21","56","41","8")
# gpsc_vec <- c("21","2","3")
gpsc_vec <- paste0("GPSC", gpsc_vec)

plot_list <- lapply(gpsc_vec, function(gpsc) {
  
  cplist <- readRDS(paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/",
           "crosspred_list_weekly_adm1_dlnm_", gpsc, "_count_bivariable_8_2019.rds"))
  
  plot_coef_crosspred(
    crosspred = cplist[[1]]$cp_high,
    type = "heatmap",
    heat_lims = c(0.78,1.22),
    exp = TRUE,
    palette = "-RdBu",
    xlab = "Lag Week",
    ylab = "PM₂.₅"
  ) + ggplot2::ggtitle(gpsc)
})

wrap_plots(plot_list)
ggsave(file = "/home/sbelman/Documents/env_sa_manuscript/figures/supplement/airpollution_GPSC_surfaces.png", width = 15, height = 10)
dev.off()
pdf(file = "/home/sbelman/Documents/env_sa_manuscript/figures/supplement/airpollution_GPSC_surfaces.pdf", width = 15, height = 10)
wrap_plots(plot_list)
dev.off()

print(plot_list[[10]] + plot_list[[9]]+plot_list[[1]] )
ggsave(file = "/home/sbelman/Documents/env_sa_manuscript/figures/supplement/airpollution_GPSC21_1_surfaces.png", width = 15, height = 5)
dev.off()

```


Calculate Numbers for Manuscripts and Tables
#### calculate N for each sub-group for Tables
```{r}
sa_adm2_weekly<-fread(file="/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm2_weekly.csv")
sa_adm1_weekly<-fread(file="/home/sbelman/Documents/env_sa_manuscript/dataframes/sa_adm1_weekly.csv")
sa_adm1_weekly %>%
  group_by(province) %>%
  summarise(disease = sum(disease),
            disease_prop = (sum(disease)/59017)*100)
sa_adm1_weekly <- sa_adm1_weekly
sa_adm1_weekly2019 <- sa_adm1_weekly[which(sa_adm1_weekly$year<2020),]

sum(sa_adm2_weekly$other_count)/sum(sa_adm2_weekly$disease)
sum(c(sa_adm2_weekly$other_count,sa_adm2_weekly$bact_count,sa_adm2_weekly$mening_count))
data2<- fread(file="/home/sbelman/Documents/env_sa_manuscript/input_datasets/disease/SA_disease_point_base.csv",quote=FALSE, header = TRUE)
dattot <- data2[which(data2$year>2004 & data2$year<2020  )]

### grids per district
shp<-st_read("/home/sbelman/Documents/env_sa_manuscript/input_datasets/shps/gadm41_namematch_ZAF_2.shp")
shp$area_km2 <- as.numeric(st_area(shp)) / 10^6
lat <- -29
grid_area_km2 <- 123 * cos(lat * pi / 180)
grid_area_km2
shp$estimated_cells <- shp$area_km2 / 107.6


# Sum each of those columns
columns_of_interest <- c("disease", "age_lt6", "age_6t14","age_15t64","age_gt65","bact_count","mening_count","other_count","nvt","pcv7","pcv13",grep("Sero",colnames(sa_adm1_weekly),value=T))  
columns_of_interest <- grep("Sero",colnames(sa_adm1_weekly),value=T)
columns_of_interest <- grep("GPSC",colnames(sa_adm1_weekly),value=T)

summary_table <- data.frame(
  sub_group = columns_of_interest,
  count = colSums(sa_adm1_weekly2019[ , ..columns_of_interest], na.rm = TRUE)
)
summary_table[order(-summary_table$count), ]

sum(sa_adm1_weekly2019$disease)
sum(sa_adm1_weekly2019$mening_count)+sum(sa_adm1_weekly2019$bact_count) + sum(sa_adm1_weekly2019$other_count)
sum(sa_adm1_weekly2019$ag)+sum(sa_adm1_weekly2019$bact_count) + sum(sa_adm1_weekly2019$other_count)

### pm2.5 and disease correlations
get_cor <- function(data){
  data %>%
    summarise(
      r = cor(pm2p5, disease, use="complete.obs"),
      n = sum(complete.cases(pm2p5, disease))
    )
}
# periods <- tibble(
#   start = c(2005,2009,2009,2005,2011,2011,2005),
#   end   = c(2009,2020,2023,2011,2023,2020,2020)
# )
periods <- list(
  "2005-2009" = c(2005,2009),
  "2009-2020" = c(2009,2020),
  "2009-2023" = c(2009,2023),
  "2005-2011" = c(2005,2011),
  "2011-2020" = c(2011,2020),
  "2011-2023" = c(2011,2023),
  "2005-2020" = c(2005,2020),
  "2005-2023" = c(2005,2023)
)
cor_table <- lapply(names(periods), function(p){

  yrs <- periods[[p]]

  sa_adm2_weekly %>%
    filter(year >= yrs[1], year <= yrs[2]) %>%
    get_cor() %>%
    mutate(period = p)

}) %>% bind_rows()
```

#### TABLE S3 - FIGURE 2 ADJACENT
#### report random effects by space and month
```{r}
endyear = 2019
space = "adm2"
time = "weekly"
if(space == "adm1"){
shp<-st_read("/home/sbelman/Documents/env_sa_manuscript/input_datasets/shps/gadm41_namematch_ZAF_1.shp")
}else{
  shp<-st_read("/home/sbelman/Documents/env_sa_manuscript/input_datasets/shps/gadm41_namematch_ZAF_2.shp")
}
N = nrow(shp$NAME_1)

### spatial model
## create a table including the random effects for weekly models at province and district level
mat_both <- NULL
for(endyears in c(2019,2023)){
  endyear <- endyears
  basemod <- readRDS(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/base_models/base_model_",time,"_20092011_popdens_",space,"_",endyear,".rds"))
  N <- length(shp$NAME_1)
  spa <- basemod$summary.random$id_u[1:N,]
  
    ## matrix
    mat_re <- matrix(nrow = N, ncol = 8)
    mat_re[,1] <- paste0("2005-",endyear)
    mat_re[,2] <- paste0("weekly_",space)
    mat_re[,3] <- "spatial"
    if (space == "adm1") {
        mat_re[,4] <- shp$NAME_1
        mat_re[,5] <- shp$NAME_1
      } else if (space == "adm2") {
        mat_re[,4] <- shp$NAME_1
        mat_re[,5] <- shp$NAME_2
      }
    mat_re[,6] <- exp(spa$mean)
    mat_re[,7] <- exp(spa$'0.025quant')
    mat_re[,8] <- exp(spa$'0.975quant')
    mat_re <- data.frame(mat_re)
    mat_re[,6:8] <- sapply(mat_re[,6:8], as.numeric)
    colnames(mat_re) <- c("years", "space_time", "random_effect","effect_indicator","effect_indicator", "RR", "RR_lower","RR_upper")
    
    ## bind
    mat_both <- rbind(mat_both,mat_re)
}


### seasonal effect
## create a table including the random effects for weekly models at province and district level
mat_both_seas <- NULL
for(endyears in c(2019,2023)){
  endyear <- endyears
  basemod <- readRDS(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/base_models/base_model_",time,"_20092011_popdens_",space,"_",endyear,".rds"))
  seasmod <- basemod$summary.random$id_m
  N <- nrow(seasmod)
  
    ## matrix
    mat_re <- matrix(nrow = N, ncol = 8)
    mat_re[,1] <- paste0("2005-",endyear)
    mat_re[,2] <- paste0("weekly_",space)
    mat_re[,3] <- "seasonal"
    mat_re[,4] <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
    mat_re[,5] <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
    mat_re[,6] <- exp(seasmod$mean)
    mat_re[,7] <- exp(seasmod$'0.025quant')
    mat_re[,8] <- exp(seasmod$'0.975quant')
    mat_re <- data.frame(mat_re)
    mat_re[,6:8] <- sapply(mat_re[,6:8], as.numeric)
    colnames(mat_re) <- c("years", "space_time", "random_effect","effect_indicator","effect_indicator", "RR", "RR_lower","RR_upper")
    
    ## bind
    mat_both_seas <- rbind(mat_both_seas,mat_re)
}
matres <- rbind(mat_both, mat_both_seas)

write.table(matres, file = paste0("/home/sbelman/Documents/env_sa_manuscript/figures/random_effects/res_",space,"_",time,".csv"), quote = FALSE, sep = ",", col.names = TRUE, row.names = FALSE)
```


## OTHER EXPLORATORY FIGURES NOT IN THE PAPER
```{r}
# read in each
  lag8 <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/nointeraction_results_fits_weekly_adm2_bivariable_8_2019_templag0humlag3.csv"))
lag8$lagtotal <- "lag8"
  lag12 <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/nointeraction_results_fits_weekly_adm2_bivariable_12_2019_templag0humlag3.csv"))
  lag12$lagtotal <- "lag12"
  lag6 <- fread(file=paste0("/home/sbelman/Documents/env_sa_manuscript/models/dlnms/bivariable/nointeraction_results_fits_weekly_adm2_bivariable_6_2019_templag0humlag3.csv"))
  lag6$lagtotal <- "lag6"
  
  ### bind together and adjust variables
  lagcomp <- rbind(lag8, lag12,lag6)
  lagcomp <- lagcomp[which(lagcomp$cov2=="absh_lag3"),]

  ## rrs
  lagcomp$rr <- exp(lagcomp$fit)
lagcomp$lowerCI_rr <- exp(lagcomp$lowerCI)
lagcomp$upperCI_rr <- exp(lagcomp$upperCI)
lagcomp$cumrr <- exp(lagcomp$cumulative_fit)
lagcomp$cumlowerCI_rr <- exp(lagcomp$cum_lowerCI)
lagcomp$cumupperCI_rr <- exp(lagcomp$cum_upperCI)

### subset for cumulative
  lagcompspec <- subset(lagcomp, lagcomp$lag_num %in% c(3))
  
  ##plot cumulative
plot_lagscum <- ggplot(lagcompspec)+
  geom_line(aes(x=var,y=cumrr,group=interaction(lag_num, lagtotal, covariate), color=lagtotal),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=cumlowerCI_rr,ymax=cumupperCI_rr, group=interaction(lag_num,lagtotal,covariate), fill=lagtotal),alpha=0.3)+
  labs(color = "", fill = "")+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10")+
  ggtitle("lag comparisons")+
  xlab("concentration")+
  facet_grid(. ~ covariate, scales = "free_x")+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "right")
plot_lagscum

plot_lagsweekly <- ggplot(lagcomp[which(lagcomp$covariate=="so2_lag0")])+
  geom_line(aes(x=var,y=rr,group=interaction(lag_num, lagtotal, covariate), color=lagtotal),alpha=0.5)+
  geom_hline(yintercept = 1, linetype = "dashed", color="red", alpha=0.6)+
  geom_ribbon(aes(x=var, ymin=lowerCI_rr,ymax=upperCI_rr, group=interaction(lag_num,lagtotal,covariate), fill=lagtotal),alpha=0.3)+
  labs(color = "", fill = "")+
  theme_bw()+
  xlab("Var")+
  ylab("Relative Risk")+
  scale_y_continuous(trans="log10")+
  ggtitle("lag comparisons")+
  xlab("concentration")+
  facet_grid(lagtotal ~ lag_num)+
  theme(axis.text = element_text(size=13),axis.title=element_text(size=13), 
        strip.text = element_text(size=13), axis.text.x = element_text(angle = 45,hjust=1, size=13),
        legend.position = "right")
plot_lagsweekly

```